### 2012-05-18 paper/paper.tex
Created skeleton of a LaTeX file with author info.

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\begin{document}

\nocite{*}

\title{Diameter and Pinched Sphere Theorems for Positively Curved Combinatorial 3-Manifolds}

\author{Aaron Trout \\ Department of Mathematics \\
Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA \and
Vadas Gintautas\\ Department of Physics \\
Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA}

\maketitle

\begin{abstract}
  Blah blah...
\end{abstract}

\section{Introduction}

\section{Basic Stuff}
\label{sect:basics}

\bibliographystyle{plain}

\bibliography{paper}

\end{document}

### 2012-05-18 paper/paper.tex
Added bibliography from old paper to new paper.

@@
 \usepackage{amsthm}                % better theorem environments
+\usepackage{url}
 
@@
 
-\title{Diameter and Pinched Sphere Theorems for Positively Curved Combinatorial 3-Manifolds}
+\title{A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds}
 
@@
 \begin{abstract}
-  Blah blah...
+  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem.
 \end{abstract}
@@
 \section{Introduction}
+
+\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and $diam(M) > \frac{\pi}{2\sqrt{\delta}}$.
+Then, $M$ is homeomorphic to a sphere.
+\end{thm}
 
@@
 
+\begin{thebibliography}{99}
+
+    \bibitem{BMM}N. Brady, J. McCammond, and J. Meier, {\em Bounding edge degrees in triangulated 3-manifolds}, Proc. Amer. Math. Soc. 132 (2004), no. 1, 291--298 (electronic).
+
+    \bibitem{cheng}Y. Cheng, {\em Eigenvalue comparison theorems and its [sic] geometric applications}, Math. Z. 143 (1975) 289-297.
+
+    \bibitem{Crowley}K. Crowley, Doctoral Thesis (Rice University), {\em Discrete Morse Theory and the Geometry of Nonpositively Curved Simplicial Complexes}. (2001).
+
+    \bibitem{EMM}M. Elder, J. McCammond, and J. Meier, {\em Combinatorial conditions that imply word-hyperbolicity for 3-manifolds.}, Topology, 42, (2003), 1241-1259.
+
+    \bibitem{forman1}R. Forman, {\em Morse theory for cell complexes}, Adv. in Math. 134 (1998), 90-145.
+
+    \bibitem{forman2}R.
+Forman, {\em Combinatorial Differential Topology and Geometry}, New perspectives in algebraic combinatorics (Berkeley, CA, 1996--97), 177--206, Math.
+Sci.
+Res.
+Inst.
+Publ., 38, Cambridge Univ.
+Press, Cambridge, 1999.
+
+    \bibitem{GGL}A.
+Gabrielov, I.
+Gelfand, and M.
+Losik.
+{\em Combinatorial computation of characteristic classes}.
+I, II, and {\em A local combinatorial formula for the first Pontrjagin class}.
+Functional Anal.
+Appl.
+9 (1975), no.
+2, 103--116; ibid.
+9 (1975), no.
+3, 186--202 (1976); ibid.
+10 (1976), no.
+1, 12--15.
+
+    \bibitem{groveshiohama}K. Grove; K, Shiohama, {\em A generalized sphere theorem}, Ann. of Math. 106 (1977), 201-211.
+
+    \bibitem{gromov}M. Gromov, {\em Curvature, diameter, and Betti numbers}, Comm. Math. Helvetici 56 (1981), 179-195.
+
+    \bibitem{LS}F. Luo and R. Stong, {\em Combinatorics of triangulations of 3-manifolds.} Trans. Amer. Math. Soc. 337 (1993), no. 2, 891-906.
+
+    \bibitem{stone}D.A. Stone, {\em Sectional curvature in piecewise linear manifolds}, Bull. Amer. Math. Soc., vol. 79 (1973), pp. 1060-1063.
+
+    \bibitem{toponogov}V.A.
+Toponogov, {\em Riemannian spaces having their curvature bounded below by a positive number}, Uspekhi Math.
+Nauk 14 (1959), 87-135 English Translation: Amer.
+Math.
+Soc.Transl.
+Ser.
+37 (1964) 287-290.
+
+    \bibitem{DDS} M. Deza, M. Dutour and M. Shtogrin, {\em On simplicial and cubical complexes with short links}, Israel Journal of Mathematics 144 (2004) 109-
+124.
+
+    \bibitem{LutzSul} F. H. Lutz and J. M. Sullivan: {\em Simplicial manifolds with small valences}.
+In preparation.
+
+\end{thebibliography}
+
 \end{document}

### 2012-05-19 paper/paper.tex
Made additional changes to Introduction (added bonnet-myers thm.)

@@
 \begin{abstract}
-  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem.
+  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem \cite{groveshiohama}.
 \end{abstract}
@@
 
+Differential geometry is of central importance not only to geometers but also topologists, physicists and, increasingly, many of those interested in applied topics such as finite element analysis REFERENCE and computer graphics REFERENCE.
+One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogs to the classical theorems and concepts from differential geometry.
+Since many way computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests REFERECE.
+Other recent DDG work can be found in \cite{BMM}, \cite{Crowley}, \cite{EMM}, \cite{forman2},
+\cite{GGL}, \cite{LS} and \cite{stone}.
+
+A particularly important goal in classical differential geometry is to elucidate the relationship between the curvature of a Riemannian (or semi-Riemannian) space and its topology. The classical
+results in this area are numerous, beautiful, and have inspired an
+enormous amount of subsequent research (REFERENCES). In this paper, we will present a discrete analog of the ``generalized'' sphere theorem of Grove and Shiohama \cite{groveshiohama}.
+
 \begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and $diam(M) > \frac{\pi}{2\sqrt{\delta}}$.
 Then, $M$ is homeomorphic to a sphere.
+\end{thm}
+
+\noindent Note that the diameter bound in this theorem is exacly half that allowed by the Bonnet-Myers theorem:
+
+\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$. Then $diam(M) \leq \frac{\pi}{\sqrt{\delta}}$.
 \end{thm}

### 2012-05-29 manifolds.py
cleaned up and renamed some of the functions in the 'functions for manifolds' section.

@@ -136,22 +136,23 @@ def pretty_print(dist):
 def get_pairs(input_list):
     return [(input_list[y],input_list[x]) for y in xrange(len(input_list)) for x in xrange(y,len(input_list)) if x!=y]
 
-#functions for manifolds.  refactor?
+#functions for manifolds.  refactor?		
 
-def traverse_edge(edge,vertex):
+def traverse_edge(edge, start_vertex):
     next_edge = list(edge)
-    next_edge.remove(vertex)
+    next_edge.remove(start_vertex)
     return next_edge[0]
 
-def get_link_path(link,next_vertex):
-    links = link[:]
-    link_path = [next_vertex]
-    for i in range(len(link)-1):
-        next_edge = [x for x in links if (next_vertex in x)][0]
-        next_vertex= traverse_edge(next_edge,next_vertex)
-        links.remove(next_edge)
-        link_path.append(next_vertex)
-    return link_path
+def get_vertices_around_circle(circle, start_vertex):
+    edges = circle[:]
+    vertices = [start_vertex]
+    next_vertex = start_vertex
+    while edges:
+        next_edge = [x for x in edges if (next_vertex in x)][0]
+        next_vertex= traverse_edge(next_edge, next_vertex)
+        edges.remove(next_edge)
+        vertices.append(next_vertex)
+    return vertices
             
 def traverse_link_path(link_path,vertex,nsteps=1):
     vertex_index = link_path.index(vertex)
@@ -160,9 +161,9 @@ def traverse_link_path(link_path,vertex,nsteps=1):
     else:
         return link_path[0]
     
-def get_opposite_link_vertex(link,edge):
-    link_path = get_link_path(link,edge[0])
-    next_vertex = traverse_link_path(link_path,edge[0])
+def get_opposite_link_vertex(link, edge):
+    link_path = get_vertices_around_circle(link, edge[0])
+    next_vertex = traverse_link_path(link_path, edge[0])
     if next_vertex in edge:
         next_vertex = traverse_link_path(link_path,next_vertex,2)
     else:
@@ -233,7 +234,7 @@ def get_jumps(manifold,vertices,edges,hops):
     #now do jumps
     jumps = []
 
-    #get edges that have diameter 5
+    #get edges that have degree 5
     link_dictionary = {}
     degree_5_edges = []
     for e in edges:
@@ -352,7 +353,6 @@ def diameter_report(graph_hash):
 
 print __name__
[…]

### 2012-05-29 manifolds.py
extended get_star and get_link to use convention that '5' is a vertex as well as [5]. So now you can just write get_star(5, manifold).

@@ -149,7 +149,7 @@ def get_vertices_around_circle(circle, start_vertex):
     next_vertex = start_vertex
     while edges:
         next_edge = [x for x in edges if (next_vertex in x)][0]
-        next_vertex= traverse_edge(next_edge, next_vertex)
+        next_vertex = traverse_edge(next_edge, next_vertex)
         edges.remove(next_edge)
         vertices.append(next_vertex)
     return vertices
@@ -172,6 +172,8 @@ def get_opposite_link_vertex(link, edge):
     return next_vertex
 
 def get_link(simplex, manifold):
+    if isinstance(simplex, (int, long)):
+		simplex = [simplex]
     link = []
     for facet in get_star(simplex, manifold):
         new_facet = facet[:]
@@ -181,6 +183,8 @@ def get_link(simplex, manifold):
     return link
 
 def get_star(simplex, manifold):
+    if isinstance(simplex, (int, long)):
+		simplex = [simplex]
     facets_in_star=[]
     for facet in manifold:
     	facet_in_star = True

### 2012-05-30 manifolds.py
reworked 'functions for manifolds' to simplify and make more general.

@@ -1,7 +1,6 @@
 import networkx, os
 import matplotlib.pyplot as plt
 import cPickle
-#import sympy
 
 edge_weight = 1
 hop_weight = (8.0/3.0)**(0.5)
@@ -147,29 +146,38 @@ def get_vertices_around_circle(circle, start_vertex):
     edges = circle[:]
     vertices = [start_vertex]
     next_vertex = start_vertex
-    while edges:
+    while len(edges) > 1:
         next_edge = [x for x in edges if (next_vertex in x)][0]
         next_vertex = traverse_edge(next_edge, next_vertex)
         edges.remove(next_edge)
         vertices.append(next_vertex)
     return vertices
-            
-def traverse_link_path(link_path,vertex,nsteps=1):
-    vertex_index = link_path.index(vertex)
-    if vertex_index < len(link_path)-nsteps:
-        return link_path[vertex_index+nsteps]
-    else:
-        return link_path[0]
-    
-def get_opposite_link_vertex(link, edge):
-    link_path = get_vertices_around_circle(link, edge[0])
-    next_vertex = traverse_link_path(link_path, edge[0])
-    if next_vertex in edge:
-        next_vertex = traverse_link_path(link_path,next_vertex,2)
-    else:
-        next_vertex = traverse_link_path(link_path,next_vertex,1)
-    
-    return next_vertex
+
+def get_opposite_simplex_in_circle(circle, simplex):
+	if isinstance(simplex, (int, long)):
+		simplex = [simplex]
+	vertices = get_vertices_around_circle(circle, simplex[0])
+	num_edges = len(circle)
+	diam = num_edges / 2
+	if len(simplex)==2:
+		other_index = vertices.index(simplex[1])
+		if num_edges % 2 == 0:
+			if other_index == (num_edges - 1):
+				return [vertices[diam-1], vertices[diam]]
+			elif other_index == 1:
+				return [vertices[diam], vertices[diam+1]]
+		if num_edges % 2 == 1:
+			if other_index == (num_edges - 1):
+				return vertices[diam]
+			elif other_index == 1:
+				return vertices[diam+1]
+	elif len(simplex)==1:
+		if num_edges % 2 == 0:
+			return vertices[diam]
+		if num_edges % 2 == 1:
+			return [vertices[diam], vertices[diam+1]]
+	else:
+		sys.exit("wrong dimension simplex")
 
 def get_link(simplex, manifold):
[…]

### 2012-05-30 manifolds.py
simplified get_pairs function implementation

@@ -110,8 +110,6 @@ class graph_hash(object):
             cPickle.dump(self.graph_dictionary,f)
             f.close()
 
-
-
 #misc functions for making things nicer
 
 def sorted_set(x):
@@ -133,7 +131,7 @@ def pretty_print(dist):
         return dist
 
 def get_pairs(input_list):
-    return [(input_list[y],input_list[x]) for y in xrange(len(input_list)) for x in xrange(y,len(input_list)) if x!=y]
+	return [(x,y) for y in input_list for x in input_list if x != y]
 
 #functions for manifolds.  refactor?		
 

### 2012-05-30 manifolds.py
improved pretty

@@ -117,18 +117,23 @@ def sorted_set(x):
 
 def pretty_print(dist):
     dist_dictionary = {}
-    for i in range(1,6):
-        dist_dictionary[i*edge_weight] = str(i)+'E'
-        dist_dictionary[i*hop_weight] = str(i)+'H'
-        dist_dictionary[i*jump_weight] = str(i)+'J'
-    dist_dictionary[edge_weight+hop_weight] = '1E+1H'
-    dist_dictionary[edge_weight+jump_weight] = '1E+1J'
-    dist_dictionary[2*edge_weight+jump_weight] = '2E+1J'
-    for d in dist_dictionary.keys():
-        if abs(d-float(dist)) < 0.000000001:
-            return dist_dictionary[d]
-    else:
-        return dist
+    max_edges = int(dist / edge_weight) + 1
+    max_hops = int(dist / hop_weight) + 1
+    max_jumps = int(dist / jump_weight) + 1
+    for edges in range(0,max_edges):
+        for hops in range(0, max_hops):
+            for jumps in range(0, max_jumps):
+                d = edges*edge_weight + hops*hop_weight + jumps*jump_weight
+                if abs(d-float(dist)) < 0.000000001:
+                    out_string = ''
+                    if (edges > 0):
+                        out_string += str(edges) + 'E '
+                    if (hops > 0):
+                        out_string += str(hops) + 'H '
+                    if (jumps > 0):
+                        out_string += str(jumps) + 'J'
+                    return out_string
+    return str(dist)
 
 def get_pairs(input_list):
 	return [(x,y) for y in input_list for x in input_list if x != y]
@@ -263,7 +268,6 @@ def get_jumps(manifold,vertices,edges,hops):
                     second_link  = link_dictionary[second_edge]
                     v1 = get_opposite_simplex_in_circle(first_link,second_edge)
                     v2 = get_opposite_simplex_in_circle(second_link,first_edge)
-                    #if (v1 != v2):
                     jumps.append(tuple(sorted([v1,v2])))
 
     return sorted_set(jumps)

### 2012-05-30 manifolds.py
cleaned up overly long lines and some other misc things

@@ -15,28 +15,30 @@ class graph_hash(object):
     debug == 1: use the first 100 manifolds as a test set
     debug == 2: use 1 simple manifold as a test
     """
-    def __init__(self,graph_dictionary_file_name,debug = 0):
+    def __init__(self, graph_dictionary_file_name, debug = 0):
 
         self.graph_dictionary_file_name = graph_dictionary_file_name
         self.debug = debug
-        self.diameter_file_name = 'diameters_' + self.graph_dictionary_file_name
+        self.diameter_file_name = 'diameters_'\
+            + self.graph_dictionary_file_name
 
         if self.debug:
             print 'Debug level:', debug
-            self.diameter_file_name = 'diameters_' + self.diameter_file_name
-            self.graph_dictionary_file_name = 'debug_' + self.graph_dictionary_file_name
+            self.diameter_file_name = 'diameters_'\
+                + self.diameter_file_name
+            self.graph_dictionary_file_name = 'debug_'\
+                + self.graph_dictionary_file_name
             self.clean()
 
         self.manifold_file_name = 'input.txt'
         self.manifold_type_file_name = 'manifolds_lex_d3_deg5.type.txt'
 
-        #initialize by building manifolds,graphs,and diameters
+        #initialize by building manifolds, graphs, and diameters
         self.manifolds = self.get_manifolds()
         self.graph_dictionary = self.load_dictionary()
 
     def clean(self):
-        #removes hashed files
-        #prompt first for safety
+        #removes hashed files, prompting first for safety
         while True:
             sure = raw_input('Delete hashed data (y/n)?')
             if 'y' in sure.lower():
@@ -52,29 +54,30 @@ class graph_hash(object):
         f = open(self.manifold_file_name,'r')
         manifolds = f.read()
         f.close()
-        #fix dumb line splitting 
-        manifolds = manifolds.replace(',\n',',').replace('\n\n','\n').replace('\n]',']') 
         
+        #fix dumb line splitting 
+        manifolds = manifolds\
+            .replace(',\n',',').replace('\n\n','\n').replace('\n]',']') 
         end = -1
         
-        if self.debug == 2: #deeper debug level -- only 1 manifold
-            manifolds = [[[1,2,3,4],[1,3,4,5],[2,3,4,6],
-                [3,4,5,7],[3,4,6,7],[3,6,7,9],
-                [4,6,7,8],[6,7,8,10],[6,7,9,10]]]
[…]

### 2012-05-30 discrete_morse.py
added new file for discrete morse function stuff

@@ -1,0 +1,32 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+#
+#  untitled.py
+#  
+#  Copyright 2012 Aaron Trout, Vadas Gintautas
+#  
+#  This program is free software; you can redistribute it and/or modify
+#  it under the terms of the GNU General Public License as published by
+#  the Free Software Foundation; either version 2 of the License, or
+#  (at your option) any later version.
+#  
+#  This program is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#  
+#  You should have received a copy of the GNU General Public License
+#  along with this program; if not, write to the Free Software
+#  Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+#  MA 02110-1301, USA.
+#  
+#  
+
+
+
+def main():
+	return 0
+
+if __name__ == '__main__':
+	main()
+

### 2012-05-30 manifolds.py

@@ -323,7 +323,7 @@ def get_weighted_graph(manifold,vertices,edges,report=False):
 
 def get_diameter(g):
     vertices = g.nodes()
-    pairs = get_pairs(vertices)
+    #pairs = get_pairs(vertices)
     dists = networkx.all_pairs_dijkstra_path_length(g)
     #for s in  shortest_paths.keys():
     #    print s,shortest_paths[s]

### 2012-05-30 discrete_morse.py
messed with discrete morse stuff

@@ -22,10 +22,28 @@
 #  
 #  
 
-
+import manifolds as mfld
 
 def main():
-	return 0
+    gh = mfld.graph_hash('manifold_graphs_small.dat', 1)
+    gd = gh.graph_dictionary
+
+    manifold_id = 23
+    base_vertex = 1
+    
+    dists = gd[manifold_id][-1]
+    m = gh.manifolds[manifold_id - 1]
+    print m
+    
+    vertices = dists.keys()
+    print vertices
+    for v in vertices:
+        d = dists[base_vertex][v]
+        lk = mfld.get_link(v, m)
+        lk_verts = mfld.get_vertices(lk)
+        ds = [w for w in lk_verts if dists[base_vertex][w] < d]
+        print ds
+    return 0
 
 if __name__ == '__main__':
 	main()

### 2012-05-30 manifolds.py
changed method of line continuation from \ to using parentheses on advice found at http://docs.python.org/howto/doanddont.html

@@ -19,15 +19,15 @@ class graph_hash(object):
 
         self.graph_dictionary_file_name = graph_dictionary_file_name
         self.debug = debug
-        self.diameter_file_name = 'diameters_'\
-            + self.graph_dictionary_file_name
+        self.diameter_file_name = ('diameters_' 
+            + self.graph_dictionary_file_name)
 
         if self.debug:
             print 'Debug level:', debug
-            self.diameter_file_name = 'diameters_'\
-                + self.diameter_file_name
-            self.graph_dictionary_file_name = 'debug_'\
-                + self.graph_dictionary_file_name
+            self.diameter_file_name = ('diameters_'
+                + self.diameter_file_name)
+            self.graph_dictionary_file_name = ('debug_'
+                + self.graph_dictionary_file_name)
             self.clean()
 
         self.manifold_file_name = 'input.txt'
@@ -97,8 +97,8 @@ class graph_hash(object):
 
     def load_dictionary(self):
         try:
-            print 'loading dictionary from '\
-                + self.graph_dictionary_file_name
+            print ('loading dictionary from '
+                + self.graph_dictionary_file_name)
             f = open(self.graph_dictionary_file_name,'rb')
             self.graph_dictionary = cPickle.load(f)
             f.close()
@@ -127,9 +127,9 @@ def pretty_print(dist):
     for num_edges in range(0,max_edges):
         for num_hops in range(0, max_hops):
             for num_jumps in range(0, max_jumps):
-                d = num_edges * edge_weight\
-                    + num_hops * hop_weight\
-                    + num_jumps * jump_weight
+                d = (num_edges * edge_weight
+                    + num_hops * hop_weight
+                    + num_jumps * jump_weight)
                 if abs(d-float(dist)) < 0.000000001:
                     out_string = ''
                     if (num_edges > 0):
@@ -212,7 +212,6 @@ def get_star(simplex, manifold):
         if facet_in_star:
     	    facets_in_star.append(facet)
     return facets_in_star
-
 
 #functions for graphs. refactor?
 
@@ -354,17 +353,17 @@ def diameter_report(graph_hash):
         out = ''
         #print t,set(type_diameter)
         for d in sorted_set(type_diameter):
-            out += str(d)[:4] + '\t' + pretty_print(d)\
-                + ' '*(9-len(pretty_print(d)))\
-                + str(type_diameter.count(d)) + '\n'
[…]

### 2012-05-31 manifolds.py
updating my latest version in order to merge with new changes

@@ -1,5 +1,5 @@
 import networkx, os
-import matplotlib.pyplot as plt
+#import matplotlib.pyplot as plt
 import cPickle
 #import sympy
 
@@ -17,6 +17,9 @@ class graph_hash(object):
     debug == 2: use 1 simple manifold as a test
     """
     def __init__(self,graph_dictionary_file_name,debug = 0):
+    """
+    initialize object
+    """
 
         self.graph_dictionary_file_name = graph_dictionary_file_name
         self.debug = debug
@@ -24,7 +27,7 @@ class graph_hash(object):
 
         if self.debug:
             print 'Debug level:', debug
-            self.diameter_file_name = 'diameters_' + self.diameter_file_name
+            self.diameter_file_name = 'debug_' + self.diameter_file_name
             self.graph_dictionary_file_name = 'debug_' + self.graph_dictionary_file_name
             self.clean()
 
@@ -109,7 +112,6 @@ class graph_hash(object):
             f = open(self.graph_dictionary_file_name,'wb')
             cPickle.dump(self.graph_dictionary,f)
             f.close()
-
 
 
 #misc functions for making things nicer
@@ -349,7 +351,6 @@ def diameter_report(graph_hash):
 
 
 
-print __name__
 if __name__ == '__main__':
 
     h = graph_hash('manifold_graphs_small.dat',1)

### 2012-05-31 manifolds.py
added GNU 2 license

@@ -1,3 +1,27 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+#
+#  manifolds.py
+#  
+#  Copyright 2012 Vadas Gintautas, Aaron Trout
+#  
+#  This program is free software; you can redistribute it and/or modify
+#  it under the terms of the GNU General Public License as published by
+#  the Free Software Foundation; either version 2 of the License, or
+#  (at your option) any later version.
+#  
+#  This program is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#  
+#  You should have received a copy of the GNU General Public License
+#  along with this program; if not, write to the Free Software
+#  Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+#  MA 02110-1301, USA.
+#  
+# 
+
 import networkx, os
 #import matplotlib.pyplot as plt
 import cPickle

### 2012-05-31 discrete_morse.py
fixed file name in comments

@@ -1,7 +1,7 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 #
-#  untitled.py
+#  discrete_morse.py
 #  
 #  Copyright 2012 Aaron Trout, Vadas Gintautas
 #  

### 2012-06-01 discrete_morse.py
messed around more with discrete_morse stuff. Also, the debug=2 example manifold actually wasn't a manifold so I replaced it with one from the list, #4784. It's kind of long in the source code, but not out of control... and it seemed like a good candidate since it has diameter 1J and non-trivial topology, L(4,1)

@@ -24,8 +24,62 @@
 
 import manifolds as mfld
 
+def get_faces(simplex):
+    if isinstance(simplex, (int, long)):
+        simplex = [simplex]
+    
+
+class SimplicialComplex:
+    '''This class represents a finite simplicial complex.
+    
+    An object of this class represents a finite simplicial complex with
+    vertices given by integers.\n'''
+    
+    def __init__(self, facets):
+        self.facets = facets
+    
+    def get_star(self, simplex):
+        if isinstance(simplex, (int, long)):
+            simplex = [simplex]
+        facets_in_star=[]
+        for facet in self.facets:
+            facet_in_star = True
+            for vertex in simplex:
+                if not vertex in facet:
+                    facet_in_star = False
+                    break
+            if facet_in_star:
+                facets_in_star.append(facet)
+        return facets_in_star
+        
+        
+class ThreeManifold(SimplicialComplex):
+    '''This class represents a compact combinatorial 3-manifold.
+    
+    An object of this class represents a compact combinatorial
+    3-manifold with vertices given by integers.\n'''
+    
+    def __init__(self, facets):
+        SimplicialComplex.__init__(self, facets)
+        # to do: code to check that link of each vertex is 2-sphere
+
 def main():
-    gh = mfld.graph_hash('manifold_graphs_small.dat', 1)
+    sc = SimplicialComplex([[1,2],[2,3,4],[3,4,5]])
+    tm = ThreeManifold([[1, 2, 3, 4], [1, 2, 3, 5], [1, 2, 4, 5],
+        [1, 3, 4, 5], [2, 3, 4, 6], [2, 3, 5, 7], [2, 3, 6, 7],
+        [2, 4, 5, 8], [2, 4, 6, 8], [2, 5, 7, 8], [2, 6, 7, 9],
+        [2, 6, 8, 9], [2, 7, 8, 9], [3, 4, 5, 10], [3, 4, 6, 10],
+        [3, 5, 7, 10], [3, 6, 7, 11], [3, 6, 10, 11], [3, 7, 10, 11],
+        [4, 5, 8, 10], [4, 6, 8, 12], [4, 6, 10, 12], [4, 8, 10, 12],
+        [5, 7, 8, 10], [6, 7, 9, 11], [6, 8, 9, 12], [6, 9, 11, 12],
[…]

### 2012-06-01 manifolds.py

@@ -62,8 +62,48 @@ class graph_hash(object):
         
         #deeper debug level -- only 1 manifold
         if self.debug == 2:
-            manifolds = [[[1,2,3,4],[1,3,4,5],[2,3,4,6], [3,4,5,7],
-                [3,4,6,7],[3,6,7,9],[4,6,7,8],[6,7,8,10],[6,7,9,10]]]
+            # The following is number 4784 in the census of positively
+            # curved manifolds (Lutz/Sullivan). It should have diameter
+            # of one jump and topological type L(4,1) 
+            manifolds = [[[1,2,3,4],[1,2,3,5],[1,2,4,6],[1,2,5,7],
+                [1,2,6,7],[1,3,4,8],[1,3,5,9],[1,3,8,9],[1,4,6,10],
+                [1,4,8,10],[1,5,7,11],[1,5,9,11],[1,6,7,12],[1,6,10,12],
+                [1,7,11,12],[1,8,9,13],[1,8,10,13],[1,9,11,13],
+                [1,10,12,13],[1,11,12,13],[2,3,4,14],[2,3,5,15],
+                [2,3,14,15],[2,4,6,16],[2,4,14,16],[2,5,7,17],
+                [2,5,15,17],[2,6,7,18],[2,6,16,18],[2,7,17,18],
+                [2,14,15,19],[2,14,16,19],[2,15,17,19],[2,16,18,19],
+                [2,17,18,19],[3,4,8,20],[3,4,14,20],[3,5,9,16],
+                [3,5,15,16],[3,8,9,21],[3,8,20,21],[3,9,16,21],
+                [3,14,15,22],[3,14,20,22],[3,15,16,22],[3,16,21,22],
+                [3,20,21,22],[4,5,6,10],[4,5,6,16],[4,5,9,11],
+                [4,5,9,16],[4,5,10,11],[4,8,10,11],[4,8,11,20],
+                [4,9,11,20],[4,9,14,16],[4,9,14,20],[5,6,10,17],
+                [5,6,15,16],[5,6,15,17],[5,7,10,11],[5,7,10,17],
+                [6,7,12,23],[6,7,18,23],[6,10,12,17],[6,12,17,23],
+                [6,15,16,18],[6,15,17,23],[6,15,18,23],[7,10,11,24],
+                [7,10,17,25],[7,10,24,25],[7,11,12,24],[7,12,23,24],
[…]

### 2012-06-01 discrete_morse.py
oops, left some code in manifolds.py I shouldn't have.

@@ -22,12 +22,14 @@
 #  
 #  
 
+import itertools
 import manifolds as mfld
-
-def get_faces(simplex):
-    if isinstance(simplex, (int, long)):
-        simplex = [simplex]
-    
+       
+def is_face_of(first_simplex, second_simplex):
+    for vertex in first_simplex:
+        if not vertex in second_simplex:
+            return False
+    return True
 
 class SimplicialComplex:
     '''This class represents a finite simplicial complex.
@@ -38,20 +40,35 @@ class SimplicialComplex:
     def __init__(self, facets):
         self.facets = facets
     
-    def get_star(self, simplex):
+    def get_closed_star(self, simplex):
         if isinstance(simplex, (int, long)):
             simplex = [simplex]
-        facets_in_star=[]
-        for facet in self.facets:
-            facet_in_star = True
-            for vertex in simplex:
-                if not vertex in facet:
-                    facet_in_star = False
-                    break
-            if facet_in_star:
-                facets_in_star.append(facet)
-        return facets_in_star
+        star = [f for f in self.facets if is_face_of(simplex, f)]
+        return SimplicialComplex(star)
         
+    def get_link(self, simplex):
+        if isinstance(simplex, (int, long)):
+            simplex = [simplex]
+        link = []
+        for facet in self.get_closed_star(simplex).facets:
+            new_facet = facet[:]
+            for vertex in simplex:
+                new_facet.remove(vertex)
+            link.append(new_facet)
+        return SimplicialComplex(link)
+        
+    def get_simplices_of_dimension(self, dim):
+        simplices = []
+        for facet in self.facets:
+            simplices += map(list, itertools.combinations(facet, dim+1))
+        return simplices
+    
+    def get_degree(self, simplex):
+        pass
+
+    def __str__(self):
+        return "Simplicial complex with facets: " + str(self.facets)
+
         
 class ThreeManifold(SimplicialComplex):
     '''This class represents a compact combinatorial 3-manifold.
[…]

### 2012-06-01 manifolds.py

@@ -132,8 +132,7 @@ class graph_hash(object):
             dists = networkx.all_pairs_dijkstra_path_length(g)
             d = get_diameter_from_dists(dists)
             self.graph_dictionary[i+1] = (g,t,d,dists)
-            if t == 'L(4,1)':
-                print (i+1), t, pretty_print(d), self.manifolds[i]
+            print (i+1), t, pretty_print(d)
         return self.graph_dictionary
 
     def load_dictionary(self):

### 2012-06-01 discrete_morse.py
added is_face_of and get_simplices_of_dimension functions

@@ -38,7 +38,7 @@ class SimplicialComplex:
     vertices given by integers.\n'''
     
     def __init__(self, facets):
-        self.facets = facets
+        self.facets = sorted(facets)
     
     def get_closed_star(self, simplex):
         if isinstance(simplex, (int, long)):
@@ -58,14 +58,12 @@ class SimplicialComplex:
         return SimplicialComplex(link)
         
     def get_simplices_of_dimension(self, dim):
-        simplices = []
+        simplices = set()
         for facet in self.facets:
-            simplices += map(list, itertools.combinations(facet, dim+1))
-        return simplices
+            s = set(itertools.combinations(facet, dim+1))
+            simplices = simplices.union(s)
+        return map(list, sorted(simplices))
     
-    def get_degree(self, simplex):
-        pass
-
     def __str__(self):
         return "Simplicial complex with facets: " + str(self.facets)
 
@@ -79,6 +77,10 @@ class ThreeManifold(SimplicialComplex):
     def __init__(self, facets):
         SimplicialComplex.__init__(self, facets)
         # to do: code to check that link of each vertex is 2-sphere
+        
+    def get_degree(self, simplex):
+        return len(self.get_closed_star(simplex).facets)
+
 
     def __str__(self):
         return "3-manifold with facets: " + str(self.facets)
@@ -97,7 +99,8 @@ def main():
     
     print tm.get_closed_star(1)
     print tm.get_link(1)
-    print tm.get_simplices_of_dimension(0)
+    print tm.get_simplices_of_dimension(2)
+    print tm.get_degree(1)
 
     
     for v in vertices:

### 2012-06-01 paper/paper.tex
changed paper to work with bibtex

--- 2012-05-18 paper/paper.tex
@@
 \begin{abstract}
-  Blah blah...
+  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem \cite{groveshiohama}.
 \end{abstract}
@@
 \section{Introduction}
+
+Differential geometry is of central importance not only to geometers but also topologists, physicists and, increasingly, many of those interested in applied topics such as finite element analysis REFERENCE and computer graphics REFERENCE.
+One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogs to the classical theorems and concepts from differential geometry.
+Since many way computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests REFERECE.
+Other recent DDG work can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,LS,stone}.
+
+A particularly important goal in classical differential geometry is to elucidate the relationship between the curvature of a Riemannian (or semi-Riemannian) space and its topology. The classical
+results in this area are numerous, beautiful, and have inspired an
+enormous amount of subsequent research (REFERENCES). In this paper, we will present a discrete analog of the ``generalized'' sphere theorem of Grove and Shiohama \cite{groveshiohama}.
+
+\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and $diam(M) > \frac{\pi}{2\sqrt{\delta}}$.
+Then, $M$ is homeomorphic to a sphere.
+\end{thm}
+
+\noindent Note that the diameter bound in this theorem is exacly half that allowed by the Bonnet-Myers theorem:
+
+\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$. Then $diam(M) \leq \frac{\pi}{\sqrt{\delta}}$.
+\end{thm}
 
@@
 \bibliographystyle{plain}
-
-\bibliography{paper}
+\bibliography{manifolds}
 

### 2012-06-03 discrete_morse.py
improved __init__ method for SimplicialComplex so that it accepts lists of simplices some of which may be faces of others. Also, added nice function found online for removing duplicates from a container. Also wrote get_descending_link function for simplicial complexes

@@ -24,6 +24,76 @@
 
 import itertools
 import manifolds as mfld
+
+## {{{ http://code.activestate.com/recipes/52560/ (r1)
+def unique(s):
+    """Return a list of the elements in s, but without duplicates.
+
+    For example, unique([1,2,3,1,2,3]) is some permutation of [1,2,3],
+    unique("abcabc") some permutation of ["a", "b", "c"], and
+    unique(([1, 2], [2, 3], [1, 2])) some permutation of
+    [[2, 3], [1, 2]].
+
+    For best speed, all sequence elements should be hashable.  Then
+    unique() will usually work in linear time.
+
+    If not possible, the sequence elements should enjoy a total
+    ordering, and if list(s).sort() doesn't raise TypeError it's
+    assumed that they do enjoy a total ordering.  Then unique() will
+    usually work in O(N*log2(N)) time.
+
+    If that's not possible either, the sequence elements must support
+    equality-testing.  Then unique() will usually work in quadratic
+    time.
+    """
+
+    n = len(s)
+    if n == 0:
+        return []
+
+    # Try using a dict first, as that's the fastest and will usually
+    # work.  If it doesn't work, it will usually fail quickly, so it
+    # usually doesn't cost much to *try* it.  It requires that all the
+    # sequence elements be hashable, and support equality comparison.
+    u = {}
+    try:
+        for x in s:
+            u[x] = 1
+    except TypeError:
+        del u  # move on to the next method
+    else:
+        return u.keys()
+
+    # We can't hash all the elements.  Second fastest is to sort,
+    # which brings the equal elements together; then duplicates are
+    # easy to weed out in a single pass.
+    # NOTE:  Python's list.sort() was designed to be efficient in the
+    # presence of many duplicate elements.  This isn't true of all
+    # sort functions in all languages or libraries, so this approach
+    # is more effective in Python than it may be elsewhere.
+    try:
+        t = list(s)
+        t.sort()
+    except TypeError:
+        del t  # move on to the next method
+    else:
+        assert n > 0
+        last = t[0]
+        lasti = i = 1
+        while i < n:
+            if t[i] != last:
+                t[lasti] = last = t[i]
+                lasti += 1
+            i += 1
+        return t[:lasti]
+
+    # Brute force is all that's left.
+    u = []
+    for x in s:
+        if x not in u:
+            u.append(x)
+    return u
[…]

### 2012-06-03 discrete_morse.py
oops, messed up some indents I think. fixed now

@@ -1,7 +1,7 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 #
-#  discrete_morse.py
+#  untitled.py
 #  
 #  Copyright 2012 Aaron Trout, Vadas Gintautas
 #  
@@ -27,25 +27,6 @@ import manifolds as mfld
 
 ## {{{ http://code.activestate.com/recipes/52560/ (r1)
 def unique(s):
-    """Return a list of the elements in s, but without duplicates.
-
-    For example, unique([1,2,3,1,2,3]) is some permutation of [1,2,3],
-    unique("abcabc") some permutation of ["a", "b", "c"], and
-    unique(([1, 2], [2, 3], [1, 2])) some permutation of
-    [[2, 3], [1, 2]].
-
-    For best speed, all sequence elements should be hashable.  Then
-    unique() will usually work in linear time.
-
-    If not possible, the sequence elements should enjoy a total
-    ordering, and if list(s).sort() doesn't raise TypeError it's
-    assumed that they do enjoy a total ordering.  Then unique() will
-    usually work in O(N*log2(N)) time.
-
-    If that's not possible either, the sequence elements must support
-    equality-testing.  Then unique() will usually work in quadratic
-    time.
-    """
 
     n = len(s)
     if n == 0:
@@ -94,6 +75,7 @@ def unique(s):
             u.append(x)
     return u
 ## end of http://code.activestate.com/recipes/52560/ }}}
+
        
 def is_face_of(first_simplex, second_simplex):
     for vertex in first_simplex:
@@ -199,6 +181,7 @@ def main():
 
     for i in range(30):
         print tm.get_descending_link(i+1, dists[1])
+  
   
 if __name__ == '__main__':
     main()

### 2012-06-03 manifolds.py

@@ -40,10 +40,8 @@ class graph_hash(object):
     debug == 2: use 1 simple manifold as a test
     """
     def __init__(self,graph_dictionary_file_name,debug = 0):
-    """
-    initialize object
-    """
-
+        """ initialize graph hash object 
+        """
         self.graph_dictionary_file_name = graph_dictionary_file_name
         self.debug = debug
         self.diameter_file_name = ('diameters_' 

### 2013-04-10 global_plot.py
added code to generate global statistics figure

@@ -1,0 +1,65 @@
+import pylab,numpy
+from matplotlib.font_manager import fontManager, FontProperties
+
+
+#set fonts here
+font= FontProperties(size='small')
+params = {'backend': 'ps',
+          'axes.labelsize': 18,
+          'xtick.labelsize': 16,
+          'ytick.labelsize': 16,
+          'legend.fontsize': 16,
+          #'font.family': 'sans',
+          'font.serif': ['Times','Palatino','serif'],
+          'font.size': 16,
+          'ps.usedistiller': 'xpdf',  
+          }
+pylab.rcParams.update(params)
+
+
+#set figure size
+csize=6 #inches = 2.95 used in chemosphere
+fig_size=(csize,csize*1.618)
+#adjust spacing
+
+
+#yt = [0, 500, 1500, 2500, 3500]  #y axis ticks
+
+fig = pylab.figure()
+
+fig.subplots_adjust(top=0.95,bottom=0.10,left=0.10,right=0.93,hspace=0.35,wspace=0.2)
+
+# this is plot 1 with Procut orange data
+
+
+ax = fig.add_subplot(1,1,1)
+
+diameters = [ 1, 1.63, 2, 2.43, 3, 3.26, 4, 4.87 ]
+steps =[ '1E', '1H', '2E', '1J', '3E', '2H', '4E', '2J' ]
+counts = [ 3, 184, 409, 2445, 1060, 582, 93, 11 ] 
+
+#ax.plot(diameters,counts,'k--',alpha=0.5)
+for d,c,s in zip(diameters,counts,steps):
+    print d,c
+    ax.plot([d,d],[1,c],'k--')
+    ax.text(d,c*1.5, s,horizontalalignment='center')
+
+ax.semilogy(diameters,counts,'ro')
+#rects1 = ax.bar(ind+width, procutMeans, width, color='r', yerr=procutStd)
+#rects1 = ax.bar(ind, procutMeans, width, color='r', yerr=procutStd)
+# add some
+ax.set_xlabel('Diameter')
+ax.set_ylabel('Count (log scale)')
+ax.set_title('Global statistics')
+#ax.set_xticks(diameters)
+#ax.set_yticks(yt)
+#ax.set_xticklabels( steps , horizontalalignment='center')
+ax.axis([0,numpy.max(diameters)*1.1,0,numpy.max(counts)*5])
+
+
[…]

### 2013-04-13 paper/paper.tex
added tables and figures to paper
Added some references and more intro material

@@
 \newtheorem{conj}[thm]{Conjecture}
+\newtheorem{dfn}[thm]{Definition}
 
@@
 
-Differential geometry is of central importance not only to geometers but also topologists, physicists and, increasingly, many of those interested in applied topics such as finite element analysis REFERENCE and computer graphics REFERENCE.
-One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogs to the classical theorems and concepts from differential geometry.
-Since many way computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests REFERECE.
-Other recent DDG work can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,LS,stone}.
+Differential geometry is of central importance not only to geometers but also topologists, physicists and, increasingly, many of those interested in applied topics such as finite element analysis and computer graphics.
+One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogues to the classical theorems and concepts from differential geometry.
+Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests \cite{grinspun2006discrete}.
+Other recent DDG work can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,stone}.
 
@@
 results in this area are numerous, beautiful, and have inspired an
-enormous amount of subsequent research (REFERENCES). In this paper, we will present a discrete analog of the ``generalized'' sphere theorem of Grove and Shiohama \cite{groveshiohama}.
+enormous amount of subsequent research. In this paper, we will present a discrete analogue of the ``generalized'' sphere theorem of Grove and Shiohama \cite{groveshiohama}.
 
@@
 Then, $M$ is homeomorphic to a sphere.
+\label{Grove-Shiohama}
 \end{thm}
 
-\noindent Note that the diameter bound in this theorem is exacly half that allowed by the Bonnet-Myers theorem:
+\noindent Note that the diameter bound in this theorem is exactly half that allowed by the Bonnet-Myers theorem:
 
@@
 
+\noindent Our results are derived from brute-force checking of a combinatorial 3-manifold census created by Lutz and Sullivan \cite{LS}.
+
 \section{Basic Stuff}
@@
 
+The discrete version of Theorem \ref{Grove-Shiohama} given in this paper will apply to positively curved combinatorial 3-manifolds.
+A {\em combinatorial 3-manifold} is an abstract simplicial complex in which the link of each vertex is a 2-sphere.
+We call such a space {\em positively curved} if at most five tetrahedra are incident along each edge.
+Why this terminology?
+If we endow a combinatorial 3-manifold with the standard piecewise-linear (PL) metric (in which all edges have unit-length), this condition is equivalent to requiring an angle deficit along each edge.
+In classical differential geometry an angle deficit is intimately related to positive curvature.
+
+Probably the most natural discrete definition of distance in an abstract simplicial complexe is given in terms of edge paths.
+
+\begin{dfn}The {\em edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum number of edges needed in an edge path from $v$ to $w$.
+We denote this by $d_1(v,w)$.
+The {\em edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$.
+\end{dfn}
+
+\noindent Previous work \cite{Trout10} has already demonstrated a discrete version of the Bonnet-Myers theorem which applies to positively curved combinatorial 3-manifolds and which uses these notions of distance and diameter.
+
+\begin{thm}[Trout] A positively curved combinatorial 3-manifold $M$ has edge-diameter at most five.
+\label{discrete_BM}
+\end{thm}
+
+While Theorem \ref{discrete_BM} is stated in terms of edge-diameter, the proof is greatly facilitated by expanding the set of paths under consideration to include paths containing not only edges, but PL-paths between vertices called {\em hops} and {\em jumps}.
+
+\begin{dfn}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$.
+We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a {\em jump} from $v_1$ to $v_2$.
+See Figure \ref{jump_fig}.
+\end{dfn}
+
+\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
+The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a {\em hop} from $v_1$ to $v_2$.
+See Figure \ref{hop_fig}.
+\end{dfn}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.6\linewidth]{figures/hops.pdf}
+        \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$, transverse to the 2-simplex $\tau$ TO DO: CROP OUT LEFT IMAGE}
+    \end{center}
+    \label{hop_fig}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=0.4\linewidth]{figures/jump.pdf}
+        \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$, transverse to the edges $e_1$ and $e_2$}
+    \end{center}
+    \label{jump_fig}
+\end{figure}
+\section{Results}
+
+diameter statistics by topological type
+
+\begin{tabular} {| l l | l l | l l | l l | l l |}
+\hline
+\multicolumn{2}{|c|}{$S^{3}/Q$} &
+\multicolumn{2}{|c|}{$L(3,1)$} &
+\multicolumn{2}{|c|}{$RP^{3}$} &
+\multicolumn{2}{|c|}{$S^{3}$} &
+\multicolumn{2}{|c|}{$L(4,1)$} \\
+\hline
+\hline
+$1H$&1    &$2E$&1    &$1H$&10    &$1E$&3    &$1J$&1 \\
+  &       &$1J$&1    &$2E$&7     &$1H$&173  &    &  \\
+  &       &  &       &$1J$&5     &$2E$&401  &    &  \\
+  &       &  &       &    &      &$1J$&2438 &    &  \\
+  &       &  &       &    &      &$3E$&1060 &    &  \\
+  &       &  &       &    &      &$2H$&582  &    &  \\
+  &       &  &       &    &      &$4E$&93   &    &  \\
+  &       &  &       &    &      &$2J$&11   &    &  \\
+\hline
+\end{tabular}
+
+global diameter statistics by topological type
+
+\begin{tabular} {| l l |}
+\hline
+$1E$ &       3\\
+$1H$ &       184\\
+$2E$ &       409\\
+$1J$ &       2445\\
+$3E$ &       1060\\
+$2H$ &       582\\
+$4E$ &       93\\
+$2J$ &       11\\
+\hline
+\end{tabular}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \caption{Global statistics}
+    \end{center}
+\end{figure}
+
 \bibliographystyle{plain}
 \bibliography{manifolds}
-
 \end{document}

### 2013-04-13 paper/paper.tex
more writing

@@
 
-\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and $diam(M) > \frac{\pi}{2\sqrt{\delta}}$.
+\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and diameter greater than $\frac{\pi}{2\sqrt{\delta}}$.
 Then, $M$ is homeomorphic to a sphere.
-\label{Grove-Shiohama}
+\label{thm:grove_shiohama}
 \end{thm}
 
-\noindent Note that the diameter bound in this theorem is exactly half that allowed by the Bonnet-Myers theorem:
+\noindent Note that this bound is sharp, since the real projective space $\mathbb{RP}^n$ admits a metric with uniform sectional curvature $K=1$ and diameter $\pi/2$.
+We should also note also that the diameter bound in Theorem \ref{thm:gove_shiohama} is exactly half the maximum diameter allowed by the Bonnet-Myers theorem:
 
-\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$. Then $diam(M) \leq \frac{\pi}{\sqrt{\delta}}$.
+\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$.
+Then the diameter of $M$ is at most $\frac{\pi}{\sqrt{\delta}}$.
+\label{thm:bonnet_myers}
 \end{thm}
@@
 
-The discrete version of Theorem \ref{Grove-Shiohama} given in this paper will apply to positively curved combinatorial 3-manifolds.
+The discrete version of Theorem \ref{thm:grove_shiohama} given in this paper will apply to positively curved combinatorial 3-manifolds.
 A {\em combinatorial 3-manifold} is an abstract simplicial complex in which the link of each vertex is a 2-sphere.
@@
 
-\begin{dfn}The {\em edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum number of edges needed in an edge path from $v$ to $w$.
-We denote this by $d_1(v,w)$.
+\begin{dfn}The {\em edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length, in edges, of an edge path from $v$ to $w$.
+We denote this quantity $d_1(v,w)$.
 The {\em edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$.
@@
 
-\noindent Previous work \cite{Trout10} has already demonstrated a discrete version of the Bonnet-Myers theorem which applies to positively curved combinatorial 3-manifolds and which uses these notions of distance and diameter.
+\noindent Previous work \cite{Trout10} has demonstrated a discrete version of the Bonnet-Myers theorem which applies to positively curved combinatorial 3-manifolds and which uses these notions of distance and diameter.
 
-\begin{thm}[Trout] A positively curved combinatorial 3-manifold $M$ has edge-diameter at most five.
-\label{discrete_BM}
+\begin{thm}[Trout] For any positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
+\label{thm:discrete_BM}
 \end{thm}
 
-While Theorem \ref{discrete_BM} is stated in terms of edge-diameter, the proof is greatly facilitated by expanding the set of paths under consideration to include paths containing not only edges, but PL-paths between vertices called {\em hops} and {\em jumps}.
+While Theorem \ref{thm:discrete_BM} is stated in terms of edge-diameter, the proof is greatly facilitated by expanding the set of paths under consideration to include those which contain not only edges, but also other types of PL-paths between vertices.
+The first new type of path is called a {\em hop}.
+
+\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
+The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a {\em hop} from $v_1$ to $v_2$.
+See Figure \ref{hop}.
+TO DO: WHY DOES THIS SAY FIGURE 2?!?!
+\end{dfn}
+
+\noindent The other new type of path we call a {\em jump.}
 
@@
 We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a {\em jump} from $v_1$ to $v_2$.
-See Figure \ref{jump_fig}.
+See Figure \ref{jump}.
 \end{dfn}
 
-\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
-The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a {\em hop} from $v_1$ to $v_2$.
-See Figure \ref{hop_fig}.
-\end{dfn}
+\noindent The length of each hop and edge will be its length as a PL-path in the standard PL metric.
+Some basic Euclidean geometry tells us these lengths are $H = \frac{2}{3}\sqrt{2}$ and $J = \frac{1}{2}\sqrt{2} + \sqrt{3}$ respectively.
+We let $d(v,w)$ denote the distance between vertices $v$ and $w$ obtained by minimizing over all paths containing edges as well as hops and jumps.
+Similarly, we let $diam(M)$ denote the diameter defined in terms of this new distance.
+
+Why are these paths relevant to this paper?
+The classical generalized sphere theorem requires the manifold to have diameter more than half the maximum allowed by the Bonnet-Myers theorem.
+If we were to naively imitate the discrete Bonnet-Myers result, Theorem \ref{thm:discrete_BM}, the corresponding diameter inequality would be $diam_1(M)>2.5$.
+Since only integral values of edge-distance can occur, this would correspond to $diam(M)\geq 3$.
+Unfortunately, this bound will not work since there are triangulations of $\mathbb{RP}^3$ with edge-diameter three.
+However, if we instead use the finer-grained measure of diameter, $diam(M)$ this problem disappears.
 
 \begin{figure}
+	\label{hop}
     \begin{center}
         \includegraphics[width=0.6\linewidth]{figures/hops.pdf}
-        \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$, transverse to the 2-simplex $\tau$ TO DO: CROP OUT LEFT IMAGE}
+        \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$. TO DO: CROP OUT LEFT IMAGE}
     \end{center}
-    \label{hop_fig}
 \end{figure}
@@
 \begin{figure}
+    \label{jump}
     \begin{center}
         \includegraphics[width=0.4\linewidth]{figures/jump.pdf}
-        \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$, transverse to the edges $e_1$ and $e_2$}
+        \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
     \end{center}
-    \label{jump_fig}
 \end{figure}
+
 \section{Results}
@@
     \end{center}
+    \label{global_statistics}
 \end{figure}

### 2013-04-14 paper/paper.tex
more writing

@@
 One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogues to the classical theorems and concepts from differential geometry.
-Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests \cite{grinspun2006discrete}.
-Other recent DDG work can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,stone}.
+Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests, see \cite{grinspun2006discrete} for example.
+Other recent DDG work with a more pure-mathematics flavor can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,stone}.
 
@@
 \noindent Note that this bound is sharp, since the real projective space $\mathbb{RP}^n$ admits a metric with uniform sectional curvature $K=1$ and diameter $\pi/2$.
-We should also note also that the diameter bound in Theorem \ref{thm:gove_shiohama} is exactly half the maximum diameter allowed by the Bonnet-Myers theorem:
+We should also note also that the diameter bound in Theorem \ref{thm:grove_shiohama} is exactly half the maximum diameter allowed by the Bonnet-Myers theorem:
 
@@
 
-Probably the most natural discrete definition of distance in an abstract simplicial complexe is given in terms of edge paths.
+Probably the most natural discrete definition of distance in an abstract simplicial complex is given in terms of edge paths.
 

### 2013-04-14 paper/paper.tex
more writing

@@
 \begin{abstract}
-  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem \cite{groveshiohama}. 
+  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem. TO DO: EXPAND ABSTRACT?
 \end{abstract}
@@
 One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogues to the classical theorems and concepts from differential geometry.
-Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests, see \cite{grinspun2006discrete} for example.
+Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests, see \cite{grinspun2006discrete} for examples.
 Other recent DDG work with a more pure-mathematics flavor can be found in \cite{BMM,Crowley,EMM,forman2,GGL1,GGL2,GGL3,stone}.
@@
 The discrete version of Theorem \ref{thm:grove_shiohama} given in this paper will apply to positively curved combinatorial 3-manifolds.
-A {\em combinatorial 3-manifold} is an abstract simplicial complex in which the link of each vertex is a 2-sphere.
-We call such a space {\em positively curved} if at most five tetrahedra are incident along each edge.
+A \textbf{combinatorial 3-manifold} $M$ is an abstract simplicial complex in which the link of each vertex is a 2-sphere.
+We call such a space \textbf{positively curved} if at most five tetrahedra are incident along each edge.
 Why this terminology?
-If we endow a combinatorial 3-manifold with the standard piecewise-linear (PL) metric (in which all edges have unit-length), this condition is equivalent to requiring an angle deficit along each edge.
+If we endow $M$ with the standard piecewise-linear (PL) metric in which all edges have unit-length, this condition is equivalent to requiring an angle deficit along each edge.
 In classical differential geometry an angle deficit is intimately related to positive curvature.
 
-Probably the most natural discrete definition of distance in an abstract simplicial complex is given in terms of edge paths.
+A very natural discrete definition of distance in an abstract simplicial complex uses \textbf{edge-paths}. That is, paths entirely contained in the 1-skeleton of $M$.
 
-\begin{dfn}The {\em edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length, in edges, of an edge path from $v$ to $w$.
+\begin{dfn}The \textbf{edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length (as a PL-path in the standard PL-metric) of an edge-path from $v$ to $w$.
 We denote this quantity $d_1(v,w)$.
-The {\em edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$.
+The \textbf{edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$.
 \end{dfn}
 
-\noindent Previous work \cite{Trout10} has demonstrated a discrete version of the Bonnet-Myers theorem which applies to positively curved combinatorial 3-manifolds and which uses these notions of distance and diameter.
+\noindent Note that the length of an edge-path is simply the number of edges it traverses.
+
+A discrete version of the Bonnet-Myers theorem was previously proved in \cite{Trout10}.
+It applies to positively curved combinatorial 3-manifolds and gives a diameter bound in terms of the edge-diameter.
 
@@
 
-While Theorem \ref{thm:discrete_BM} is stated in terms of edge-diameter, the proof is greatly facilitated by expanding the set of paths under consideration to include those which contain not only edges, but also other types of PL-paths between vertices.
+\noindent Interestingly, while the statement of Theorem \ref{thm:discrete_BM} uses edge-diameter, its proof relies on expanding the set of paths under consideration to include those which contain not only edges, but also other types of PL-paths between vertices.
+
+\subsection{Hops and Jumps}
 The first new type of path is called a {\em hop}.
 
-\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
-The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a {\em hop} from $v_1$ to $v_2$.
-See Figure \ref{hop}.
-TO DO: WHY DOES THIS SAY FIGURE 2?!?!
+\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ are vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
+The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a \textbf{hop} from $v_1$ to $v_2$.
+See Figure \ref{fig:hop}.
+TO DO: FIX THIS FIG NUM
 \end{dfn}
 
-\noindent The other new type of path we call a {\em jump.}
-
-\begin{dfn}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$.
-We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a {\em jump} from $v_1$ to $v_2$.
-See Figure \ref{jump}.
-\end{dfn}
-
-\noindent The length of each hop and edge will be its length as a PL-path in the standard PL metric.
-Some basic Euclidean geometry tells us these lengths are $H = \frac{2}{3}\sqrt{2}$ and $J = \frac{1}{2}\sqrt{2} + \sqrt{3}$ respectively.
-We let $d(v,w)$ denote the distance between vertices $v$ and $w$ obtained by minimizing over all paths containing edges as well as hops and jumps.
-Similarly, we let $diam(M)$ denote the diameter defined in terms of this new distance.
-
-Why are these paths relevant to this paper?
-The classical generalized sphere theorem requires the manifold to have diameter more than half the maximum allowed by the Bonnet-Myers theorem.
-If we were to naively imitate the discrete Bonnet-Myers result, Theorem \ref{thm:discrete_BM}, the corresponding diameter inequality would be $diam_1(M)>2.5$.
-Since only integral values of edge-distance can occur, this would correspond to $diam(M)\geq 3$.
-Unfortunately, this bound will not work since there are triangulations of $\mathbb{RP}^3$ with edge-diameter three.
-However, if we instead use the finer-grained measure of diameter, $diam(M)$ this problem disappears.
-
 \begin{figure}
-	\label{hop}
+	\label{fig:hop}
     \begin{center}
@@
 
+\noindent The other new type of path we call a {\em jump.}
+
+\begin{dfn}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$.
+We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a \textbf{jump} from $v_1$ to $v_2$.
+See Figure \ref{fig:jump}.
+TO DO: FIX THIS FIG NUM
+\end{dfn}
+
 \begin{figure}
-    \label{jump}
+    \label{fig:jump}
     \begin{center}
@@
 
+\noindent Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric.
+Some basic Euclidean geometry tells us these lengths are $H = \frac{2}{3}\sqrt{2}$ and $J = \frac{1}{2}\sqrt{2} + \sqrt{3}$ respectively.
+We let $d(v,w)$ denote the \textbf{distance between vertices $v$ and $w$} obtained by minimizing over all paths containing edges as well as hops and jumps.
+Similarly, we let $diam(M)$ denote the \textbf{diameter of $M$} defined in terms of the distance function $d$.
+
+Why are these paths relevant to this paper?
+The classical generalized sphere theorem requires the manifold to have diameter more than half the maximum allowed by the classical Bonnet-Myers theorem.
+If we were to naively imitate the classical results in the discrete setting, from Theorem \ref{thm:discrete_BM} we would want the bound $diam_1(M)>\frac{5}{2}$.
+Since only integral values of edge-distance can occur, this would correspond to $diam_1(M)\geq 3$.
+Unfortunately, there are positively curved combinatorial 3-manifolds with edge-diameter three that are not homeomorphic to the 3-sphere.
+In particular, there exists positively curved triangulations of $\mathbb{RP}^3$ with edge-diameter three.
+However, if we instead use the finer-grained measure of diameter, $diam(M)$ this problem disappears and the discrete results again mirror the classical ones.
+
+From the proof of Theorem \ref{thm:discrete_BM} in \cite{Trout10}, we have the following bound on the possible distances which occur in a positively curved 3-manifold.
+
+\begin{lem}[Trout] If $v$ and $w$ are vertices in a positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
+\end{lem}
+
+\noindent Note that we have listed the possible distances in increasing numerical order. This result implies a diameter bound, which is known to be sharp by results in \cite{Trout10}.
+
+\begin{thm} If $M$ is a positively curved combinatorial 3-manifold then $diam(M) \leq 2J$.
+\label{thm:discrete_BM_expanded_paths}
+\end{thm}
+
+\noindent This is essentially Theorem \ref{thm:discrete_BM} stated in terms of our new expanded set of paths. It is this result which we will consider our ``discrete Bonnet-Myers'' theorem.
+
 \section{Results}
+We are now finally able to state our main result, which shows that any positively curved 3-manifold with diameter larger than half that allowed by Corollary \ref{thm:discrete_BM_expanded_paths}.
+
+\begin{thm} Any positively curved combinatorial 3-manifold $M$ with $diam(M)>J$ is homeomorphic to the 3-sphere.
+\label{thm:discrete_GS}
+\end{thm}
 

### 2013-04-14 paper/paper.tex
more writing

@@
 
-\section{Results}
+\section{Main Results}
+
 We are now finally able to state our main result, which shows that any positively curved 3-manifold with diameter larger than half that allowed by Corollary \ref{thm:discrete_BM_expanded_paths}.
@@
 \begin{thm} Any positively curved combinatorial 3-manifold $M$ with $diam(M)>J$ is homeomorphic to the 3-sphere.
+This bound is sharp, and equal to half the maximum diameter which occurs for any such $M$.
 \label{thm:discrete_GS}
@@
 
-diameter statistics by topological type
+\noindent The proof of this result relies on a complete census of positively curved combinatorial 3-manifolds completed by Lutz and Sullivan.
+The census contains 4787 manifolds and contains examples of the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$ and the cube space $S^3/Q$.
+The creation of such a census is a highly non-trivial task, for details see \cite{LutzSul}.
+
+Here is a table of the number of triangulations in the census at each topological type and diameter.\vspace{.1in}
 
@@
 \hline
-\end{tabular}
+\end{tabular}\vspace{.1in}
 
-global diameter statistics by topological type
+\noindent Here is a table of the number of triangulations at each diameter. \vspace{.1in}
 
@@
 
+\section{Algorithm Used}
+
 \bibliographystyle{plain}

### 2013-04-14 paper/paper.tex
more writing

@@
 
-\section{Basic Stuff}
+\section{Preliminaries}
 \label{sect:basics}
@@
 
-We are now finally able to state our main result, which shows that any positively curved 3-manifold with diameter larger than half that allowed by Corollary \ref{thm:discrete_BM_expanded_paths}.
+We are now able to state our main result.
 
@@
 
-\noindent The proof of this result relies on a complete census of positively curved combinatorial 3-manifolds completed by Lutz and Sullivan.
-The census contains 4787 manifolds and contains examples of the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$ and the cube space $S^3/Q$.
+\noindent The proof of this theorem relies on a complete census of positively curved combinatorial 3-manifolds completed by Lutz and Sullivan.
+The census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$ and the cube space $S^3/Q$.
 The creation of such a census is a highly non-trivial task, for details see \cite{LutzSul}.
@@
     \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
-    \caption{Global statistics}
+    \caption{Global statistics.
+TO DO: Change diameter labels to $1, H, 2, J, 3, 2H, 4, 2J$ and put these on horizontal axis.
+TO DO: Change graph to show separate data points for each topological type.
+}
     \end{center}
@@
 
+TO DO: Give a brief description of the algorithm used.
+
+\section{Discussion}
+
+TO DO: Discuss the big picture stuff concerning adding possible PL paths between vertices. Mention the benefits of hops and jumps not mentioned here (unique continuation of minimal paths)
+
+We should mention another notable advantage of expanding the space of paths to include hops and edges.
+In classical differential geometry, the starting point, the initial direction, and length completely determine a geodesic.
+There is no branching of geodesics.
+This property of fundamental importance to many proofs in Riemannian geometry.
+If we use only edge-paths, then in a positively curved combinatorial 3-manifold it is possible to have two minimal paths with the same length and which begin by traversing the same edge, yet reach distinct destination vertices.
+That is, using only edge paths, minimal geodesics can ``branch''.
+
+This important property is restored if we include hops and edges. In \cite{Trout10} the following is shown.
+
+\begin{thm}
+\end{thm}
+
 \bibliographystyle{plain}

### 2013-04-15 paper/paper.tex
more writing

@@
 
-\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$.
+\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with sectional curvature $K \geq \delta > 0$.
 Then the diameter of $M$ is at most $\frac{\pi}{\sqrt{\delta}}$.
@@
 
-A very natural discrete definition of distance in an abstract simplicial complex uses \textbf{edge-paths}. That is, paths entirely contained in the 1-skeleton of $M$.
+A very natural discrete definition of distance in an abstract simplicial complex uses \textbf{edge-paths}.
+That is, paths entirely contained in the 1-skeleton of $M$.
+Let $\mathcal{P}_1$ denote the set of all edge-paths on $M$.
 
@@
 \noindent Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric.
-Some basic Euclidean geometry tells us these lengths are $H = \frac{2}{3}\sqrt{2}$ and $J = \frac{1}{2}\sqrt{2} + \sqrt{3}$ respectively.
+Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:jump_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively.
 We let $d(v,w)$ denote the \textbf{distance between vertices $v$ and $w$} obtained by minimizing over all paths containing edges as well as hops and jumps.
 Similarly, we let $diam(M)$ denote the \textbf{diameter of $M$} defined in terms of the distance function $d$.
+Let $\mathcal{P}$ denote the set of all the paths containing edges, hops or jumps.
 
 Why are these paths relevant to this paper?
-The classical generalized sphere theorem requires the manifold to have diameter more than half the maximum allowed by the classical Bonnet-Myers theorem.
-If we were to naively imitate the classical results in the discrete setting, from Theorem \ref{thm:discrete_BM} we would want the bound $diam_1(M)>\frac{5}{2}$.
-Since only integral values of edge-distance can occur, this would correspond to $diam_1(M)\geq 3$.
+In the classical setting, the generalized sphere theorem requires a manifold to have diameter more than half the maximum allowed by the Bonnet-Myers theorem, and this diameter bound cannot be improved.
+If we naively imitate the classical results in the discrete setting, from Theorem \ref{thm:discrete_BM} we would want to assume the bound $diam_1(M)>\frac{5}{2}$.
 Unfortunately, there are positively curved combinatorial 3-manifolds with edge-diameter three that are not homeomorphic to the 3-sphere.
-In particular, there exists positively curved triangulations of $\mathbb{RP}^3$ with edge-diameter three.
-However, if we instead use the finer-grained measure of diameter, $diam(M)$ this problem disappears and the discrete results again mirror the classical ones.
+In particular, as we will see, $\mathbb{RP}^3$ has positively curved triangulations with edge-diameter three.
+However, if we instead use the finer-grained measure of diameter, $diam(M)$, the problem disappears and the discrete results mirror the classical ones exactly.
+That is, we can prove a discrete version of the generalized sphere theorem with a diameter bound exactly half the maximum allowed by the corresponding discrete Bonnet-Myers theorem.
 
+What is the discrete Bonnet-Myers theorem corresponding to Theorem \ref{thm:discrete_BM} but which uses $diam(M)$?
 From the proof of Theorem \ref{thm:discrete_BM} in \cite{Trout10}, we have the following bound on the possible distances which occur in a positively curved 3-manifold.
@@
 
-\noindent Note that we have listed the possible distances in increasing numerical order. This result implies a diameter bound, which is known to be sharp by results in \cite{Trout10}.
+\noindent Note that we have listed the possible distances in increasing numerical order. This result implies the desired discrete Bonnet-Myers type theorem.
 
@@
 
-\noindent This is essentially Theorem \ref{thm:discrete_BM} stated in terms of our new expanded set of paths. It is this result which we will consider our ``discrete Bonnet-Myers'' theorem.
+\noindent Note, it is shown in \cite{Trout10} that this diameter bound is sharp.
 
@@
 \noindent The proof of this theorem relies on a complete census of positively curved combinatorial 3-manifolds completed by Lutz and Sullivan.
-The census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$ and the cube space $S^3/Q$.
+The census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
 The creation of such a census is a highly non-trivial task, for details see \cite{LutzSul}.
@@
 
-This important property is restored if we include hops and edges. In \cite{Trout10} the following is shown.
+This important property is restored if we include hops and jumps. In \cite{Trout10} the following is shown.
 
-\begin{thm}
+\begin{thm}(Unique Extension) If two paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of $M$ and have the same length, then $P_1 = P_2$.
+This result is false if $\mathcal{P}$ is replaced by $\mathcal{P}_1$.
 \end{thm}

### 2013-04-15 paper/paper.tex
more writing

@@
 
-\noindent The proof of this theorem relies on a complete census of positively curved combinatorial 3-manifolds completed by Lutz and Sullivan.
-The census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
-The creation of such a census is a highly non-trivial task, for details see \cite{LutzSul}.
+\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds (see \cite{Lutz_online}.) This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
+Note that all manifolds are either homeomorphic to a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.
 
@@
 
-\begin{thm}(Unique Extension) If two paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of $M$ and have the same length, then $P_1 = P_2$.
+\begin{thm}(Unique Extension) If two minimal paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of a positively curved 3-manifold $M$ and $P_1$ and $P_2$ have the same length, then $P_1 = P_2$.
 This result is false if $\mathcal{P}$ is replaced by $\mathcal{P}_1$.

### 2013-04-15 paper/paper.tex
more writing, messing with figures, more to-dos

@@
 
-\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds (see \cite{Lutz_online}.) This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
+\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types, see \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}.
+This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
 Note that all manifolds are either homeomorphic to a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.
 
-Here is a table of the number of triangulations in the census at each topological type and diameter.\vspace{.1in}
+Here is a table of the number of triangulations in the census at each topological type and diameter.
+TO DO: CHANGE ORDER OF MANIFOLDS IN TABLE TO S3/Q, L(4,1), L(3,1), RP3, S3.
+TO DO: PUT EACH POSSIBLE DIAMETER IN TABLE ON ITS OWN COLUMN (TO ALIGN THEM WITH LABELS ON THE LEFT...) \vspace{.1in}
 
@@
 \hline
-\end{tabular}\vspace{.1in}
+\end{tabular}
 
-\noindent Here is a table of the number of triangulations at each diameter. \vspace{.1in}
+\noindent Here is a table of the number of triangulations at each diameter. TO DO: MAKE THIS A "TOTAL" COLUMN IN PREVIOUS TABLE.\vspace{.1in}
 
@@
     \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \includegraphics[width=0.9\linewidth]{figures/global_statistics.png}
     \caption{Global statistics.

### 2013-04-15 paper/paper.tex
misc stuff

@@
     \caption{Global statistics.
-TO DO: Change diameter labels to $1, H, 2, J, 3, 2H, 4, 2J$ and put these on horizontal axis.
-TO DO: Change graph to show separate data points for each topological type.
-}
+TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS.
+TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE.
+TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS, WHERE THE CORRESPONDENCE BETWEEN DIAMETER(s) AND EDGE-DIAMETER IS $0\rightarrow 0, 1\rightarrow 1, \{H, 2\}\rightarrow 2, \{J,3\}\rightarrow 3, 4\rightarrow 4, 2J\rightarrow 5$}
     \end{center}
@@
 
-TO DO: Give a brief description of the algorithm used.
+TO DO: BRIEF DISCUSSION OF ALGORITHM USED
 
@@
 
-TO DO: Discuss the big picture stuff concerning adding possible PL paths between vertices. Mention the benefits of hops and jumps not mentioned here (unique continuation of minimal paths)
+TO DO: BIG PICTURE STUFF ABOUT EXTENDIND EDGE PATHS. WHY DOES IT HELP? CAN WE USE IT TO DO OTHER NICE THINGS? INCORPORATE FOLLOWING STUFF INTO DISCUSSION ...
 
@@
 This property of fundamental importance to many proofs in Riemannian geometry.
-If we use only edge-paths, then in a positively curved combinatorial 3-manifold it is possible to have two minimal paths with the same length and which begin by traversing the same edge, yet reach distinct destination vertices.
+If we use only edge-paths, then in a positi`vely curved combinatorial 3-manifold it is possible to have two minimal paths with the same length and which begin by traversing the same edge, yet reach distinct destination vertices.
 That is, using only edge paths, minimal geodesics can ``branch''.

### 2013-06-03 paper/paper.tex
fixed table

@@
 See Figure \ref{fig:hop}.
-TO DO: FIX THIS FIG NUM
+TO DO: FIX THIS FIG NUM  VADAS: I THINK THIS IS DUE TO AMSTHM AND I``M NOT SURE HOW TO FIX IT.
 \end{dfn}
@@
 
-Here is a table of the number of triangulations in the census at each topological type and diameter.
-TO DO: CHANGE ORDER OF MANIFOLDS IN TABLE TO S3/Q, L(4,1), L(3,1), RP3, S3.
-TO DO: PUT EACH POSSIBLE DIAMETER IN TABLE ON ITS OWN COLUMN (TO ALIGN THEM WITH LABELS ON THE LEFT...) \vspace{.1in}
+Here is a table of the number of triangulations in the census at each topological type and diameter. \vspace{.1in}
 
-\begin{tabular} {| l l | l l | l l | l l | l l |}
+\begin{tabular} {| c | c | c | c | c | c | c |}
     \hline
-\multicolumn{2}{|c|}{$S^{3}/Q$} &
-\multicolumn{2}{|c|}{$L(3,1)$} &
-\multicolumn{2}{|c|}{$RP^{3}$} &
-\multicolumn{2}{|c|}{$S^{3}$} &
-\multicolumn{2}{|c|}{$L(4,1)$} \\
+Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
 \hline
 \hline
-$1H$&1    &$2E$&1    &$1H$&10    &$1E$&3    &$1J$&1 \\
-  &       &$1J$&1    &$2E$&7     &$1H$&173  &    &  \\
-  &       &  &       &$1J$&5     &$2E$&401  &    &  \\
-  &       &  &       &    &      &$1J$&2438 &    &  \\
-  &       &  &       &    &      &$3E$&1060 &    &  \\
-  &       &  &       &    &      &$2H$&582  &    &  \\
-  &       &  &       &    &      &$4E$&93   &    &  \\
-  &       &  &       &    &      &$2J$&11   &    &  \\
+$1E$&  &  &  &   &3    & 3    \\
+$1H$&1 &  &  &10 &173  & 184  \\
+$2E$&  &  &1 &7  &401  & 409  \\
+$1J$&  &1 &1 &5  &2438 & 2445 \\
+$3E$&  &  &  &   &1060 & 1060 \\
+$2H$&  &  &  &   &582  & 582  \\
+$4E$&  &  &  &   &93   & 93   \\
+$2J$&  &  &  &   &11   & 11   \\
+\hline
+Total&1 &1 &2 &22   &4761   &4787  \\
 \hline
@@
 
-\noindent Here is a table of the number of triangulations at each diameter. TO DO: MAKE THIS A "TOTAL" COLUMN IN PREVIOUS TABLE.\vspace{.1in}
-
-\begin{tabular} {| l l |}
-\hline
-$1E$ &       3\\
-$1H$ &       184\\
-$2E$ &       409\\
-$1J$ &       2445\\
-$3E$ &       1060\\
-$2H$ &       582\\
-$4E$ &       93\\
-$2J$ &       11\\
-\hline
-\end{tabular}
+\noindent Here is a table of the number of triangulations at each diameter.
 

### 2013-06-10 paper/paper.tex
more writing. expanded abstract, fixed some figs

@@
 \begin{abstract}
-  This paper presents a discrete version of Grove and Shiohama's generalized sphere theorem. TO DO: EXPAND ABSTRACT?
+The generalized sphere theorem of Grove and Shiohama is a central result in the differential geometry of positively curved manifolds.
+It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere.
+In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds.
+That is, those for which at most five tetrahedra are incident at each edge.
+In the standard piecewise-linear metric, this bound is equivalent to requiring an angle deficit along each edge.
+Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem.
+Here we prove a corresponding {\em discrete} generalized sphere theorem using a complete census of positively curved 3-manifolds completed by Lutz and Sullivan.
 \end{abstract}
@@
 
-\noindent Our results are derived from brute-force checking of a combinatorial 3-manifold census created by Lutz and Sullivan \cite{LS}.
+\noindent The main result of this paper is a discrete version of Theorem \ref{thm:grove_shiohama} which is proved by brute-force checking of a combinatorial 3-manifold census created by Lutz and Sullivan \cite{LS}.
+Before we can state precisely this result, we need a few preliminaries.
 
@@
 
-\begin{thm}[Trout] For any positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
+\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
 \label{thm:discrete_BM}
@@
 The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a \textbf{hop} from $v_1$ to $v_2$.
-See Figure \ref{fig:hop}.
-TO DO: FIX THIS FIG NUM  VADAS: I THINK THIS IS DUE TO AMSTHM AND I``M NOT SURE HOW TO FIX IT.
+See Figure \ref{fig:hop} for an illustration.
 \end{dfn}
@@
 \begin{figure}
+	\centering
+		\includegraphics[width=0.23\linewidth]{figures/hops.pdf}
+    \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$.}
     	\label{fig:hop}
-    \begin{center}
-        \includegraphics[width=0.6\linewidth]{figures/hops.pdf}
-        \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$. TO DO: CROP OUT LEFT IMAGE}
-    \end{center}
 \end{figure}
@@
 We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a \textbf{jump} from $v_1$ to $v_2$.
-See Figure \ref{fig:jump}.
-TO DO: FIX THIS FIG NUM
+For an illustration, see Figure \ref{fig:jump}.
 \end{dfn}
@@
 \begin{figure}
+	\centering
+		\includegraphics[width=0.30\linewidth]{figures/jump.pdf}
+    \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
     \label{fig:jump}
-    \begin{center}
-        \includegraphics[width=0.4\linewidth]{figures/jump.pdf}
-        \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
-    \end{center}
 \end{figure}
 
-\noindent Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric.
+Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric.
 Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:jump_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively.
@@
 
-\begin{thm} If $M$ is a positively curved combinatorial 3-manifold then $diam(M) \leq 2J$.
+\begin{thm}[Trout] For any positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
 \label{thm:discrete_BM_expanded_paths}
@@
 
-\begin{thm} Any positively curved combinatorial 3-manifold $M$ with $diam(M)>J$ is homeomorphic to the 3-sphere.
+\begin{thm} Any connected positively curved combinatorial 3-manifold $M$ with $diam(M)>J$ is homeomorphic to the 3-sphere.
 This bound is sharp, and equal to half the maximum diameter which occurs for any such $M$.
@@
 
-Here is a table of the number of triangulations in the census at each topological type and diameter. \vspace{.1in}
+Here is a table of the number of triangulations in the census at each topological type and diameter.
 
-\begin{tabular} {| c | c | c | c | c | c | c |}
+\begin{figure}
+\centering
+\begin{tabular} {| r | r | r | r | r | r | r |}
 \hline
@@
 \hline
-$1E$&  &  &  &   &3    & 3    \\
-$1H$&1 &  &  &10 &173  & 184  \\
-$2E$&  &  &1 &7  &401  & 409  \\
-$1J$&  &1 &1 &5  &2438 & 2445 \\
-$3E$&  &  &  &   &1060 & 1060 \\
+$1$&  &  &  &   &3    & 3    \\
+$H$&1 &  &  &10 &173  & 184  \\
+$2$&  &  &1 &7  &401  & 409  \\
+$J$&  &1 &1 &5  &2438 & 2445 \\
+$3$&  &  &  &   &1060 & 1060 \\
 $2H$&  &  &  &   &582  & 582  \\
-$4E$&  &  &  &   &93   & 93   \\
+$4$&  &  &  &   &93   & 93   \\
 $2J$&  &  &  &   &11   & 11   \\
@@
 \end{tabular}
+\caption{The positively curved combinatorial 3-manifolds by diameter and topological type}
+\label{fig:manifold_counts}
+\end{figure}
 
-\noindent Here is a table of the number of triangulations at each diameter.
+\noindent Compare this table with the corresponding counts indexed by the coarser {\em edge-diameter}.
+
+\begin{figure}
+\centering
+\begin{tabular} {| r | r | r | r | r | r | r |}
+\hline
+Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
+\hline
+\hline
+$1$ &  &  &  &   &3    & 3    \\
+$2$ & 1&  &1 &17 &574  & 593  \\
+$3$ &  &1 &1 &5  &3498 & 3505 \\
+$4$ &  &  &  &   &675  & 675   \\
+$5$ &  &  &  &   &11   & 11   \\
+\hline
+Total&1 &1 &2 &22   &4761   &4787  \\
+\hline
+\end{tabular}
+\caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
+\label{fig:manifold_counts}
+\end{figure}
 
@@
     \begin{center}
-    \includegraphics[width=0.9\linewidth]{figures/global_statistics.png}
+    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
     \caption{Global statistics.
@@
 
-This important property is restored if we include hops and jumps. In \cite{Trout10} the following is shown.
+This important property is restored, at least for positively curved manifolds, if we include hops and jumps. In \cite{Trout10} the following is shown.
 

### 2013-06-10 paper/paper.tex
more writing.

@@
 
-\begin{lem}[Trout] If $v$ and $w$ are vertices in a positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
+\begin{lem}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
 \end{lem}
@@
 
-\begin{thm}[Trout] For any positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
+\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
 \label{thm:discrete_BM_expanded_paths}
@@
 This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
-Note that all manifolds are either homeomorphic to a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.
+Note that all of these manifolds are homeomorphic to either a 3-sphere or to the quotient space of a 3-sphere under the action of a finite group.
 
-Here is a table of the number of triangulations in the census at each topological type and diameter.
+A small Python program was created to examine each manifold in the census and compute its diameter.
+See Figure \ref{fig:global_statistics} for a plot of the number of manifolds by topological type and diameter.
+Note that, as claimed, every manifold with diameter greater than $J$ is a 3-sphere.
+The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:global_statistics_edges}.
 
@@
 
-\noindent Compare this table with the corresponding counts indexed by the coarser {\em edge-diameter}.
-
 \begin{figure}
@@
 \caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
-\label{fig:manifold_counts}
+\label{fig:manifold_counts_edges}
 \end{figure}
@@
 TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE.
-TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS, WHERE THE CORRESPONDENCE BETWEEN DIAMETER(s) AND EDGE-DIAMETER IS $0\rightarrow 0, 1\rightarrow 1, \{H, 2\}\rightarrow 2, \{J,3\}\rightarrow 3, 4\rightarrow 4, 2J\rightarrow 5$}
+TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS}
     \end{center}
-    \label{global_statistics}
+    \label{fig:global_statistics}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \caption{...TO DO}
+    \end{center}
+    \label{fig:global_statistics_edges}
 \end{figure}

### 2013-06-10 paper/paper.tex
more writing.

@@
 See Figure \ref{fig:global_statistics} for a plot of the number of manifolds by topological type and diameter.
-Note that, as claimed, every manifold with diameter greater than $J$ is a 3-sphere.
+As claimed, every manifold with diameter greater than $J$ is a 3-sphere.
 The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:global_statistics_edges}.
+Note that there are $3$-manifolds homeomor
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \caption{Global statistics.
+TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS.
+TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE.
+TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS}
+    \end{center}
+    \caption{TO DO: ADD CAPTION}
+    \label{fig:global_statistics}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \end{center}
+    \caption{TO DO: ADD CAPTION}
+    \label{fig:global_statistics_edges}
+\end{figure}
 
@@
 
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
-    \caption{Global statistics.
-TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS.
-TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE.
-TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS}
-    \end{center}
-    \label{fig:global_statistics}
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
-    \caption{...TO DO}
-    \end{center}
-    \label{fig:global_statistics_edges}
-\end{figure}
-
 \section{Algorithm Used}
@@
 \section{Discussion}
-
-TO DO: BIG PICTURE STUFF ABOUT EXTENDIND EDGE PATHS. WHY DOES IT HELP? CAN WE USE IT TO DO OTHER NICE THINGS? INCORPORATE FOLLOWING STUFF INTO DISCUSSION ...
+Why does adding additional paths above and beyond edge-paths restore the classical structure to our proof?
 

### 2013-06-11 paper/paper.tex
updated graphs per TO DOs

@@
     \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
-    \caption{Global statistics.
-TO DO: CHANGE DIAMETER LABELS TO $1, H, 2, J, 3, 2H, 4, 2J$ AND PUT THESE ON HORIZONTAL AXIS.
-TO DO: CHANGE GRAPH TO SHOW SEPARATE DATA POINTS FOR EACH TOPOLOGICAL TYPE.
-TO DO: CREATE ANOTHER GRAPH WITH JUST EDGE DIAMETER COUNTS}
     \end{center}
-    \caption{TO DO: ADD CAPTION}
+    \caption{Distribution of diameters.}
     \label{fig:global_statistics}
@@
     \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
+    \includegraphics[width=0.6\linewidth]{figures/split_statistics.png}
     \end{center}
-    \caption{TO DO: ADD CAPTION}
+    \caption{Distribution of diameters for each type.}
+    \label{fig:type_statistics}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.6\linewidth]{figures/edge_diameters.png}
+    \end{center}
+    \caption{Distribution of diameters when only edges are considered, without hops or jumps.}
     \label{fig:global_statistics_edges}

### 2013-06-11 edge_plot.py

@@ -34,36 +34,24 @@ fig.subplots_adjust(top=0.95,bottom=0.1,left=0.10,right=0.93,hspace=0.35,wspace=
 
 ax = fig.add_subplot(1,1,1)
 
-diameters = [ 1, 1.63, 2, 2.43, 3, 3.26, 4, 4.87 ]
-steps =[ '1', 'H', '2', 'J', '3', '2H', '4', '2J' ]
-counts = [ 3, 184, 409, 2445, 1060, 582, 93, 11 ] 
-
-diameters1= [ 1, 2, 3, 4, ]
-counts1= [ 3, 409, 1060, 93 ] 
-
-diameters2 = [ 1.63, 2.43, 3.26, 4.87 ]
-steps2 =[ 'H', 'J', '2H', '2J' ]
-counts2 = [ 184, 2445, 582, 11 ] 
+diameters = [ 1, 2, 3, 4, 5 ]
+counts = [ 3, 593,3505, 675, 11 ] 
 
 #ax.plot(diameters,counts,'k--',alpha=0.5)
-for d,c in zip(diameters1,counts1):
+for d,c in zip(diameters,counts):
     ax.plot([d,d],[1,c],'k--')
-
-for d,c,s in zip(diameters2,counts2,steps2):
-    ax.plot([d,d],[2.3,c],'k--')
-    ax.text(d,1.3,s,horizontalalignment='center')
 
 ax.semilogy(diameters,counts,'ro')
 #rects1 = ax.bar(ind+width, procutMeans, width, color='r', yerr=procutStd)
 #rects1 = ax.bar(ind, procutMeans, width, color='r', yerr=procutStd)
 # add some
-ax.set_xlabel('Diameter')
+ax.set_xlabel('Diameter with only Edges')
 #ax.xaxis.set_label_coords(0.5, -0.12)
 ax.set_ylabel('Count (log scale)')
 ax.set_title('Global statistics')
-ax.set_xticks([0] + diameters + [5,6])
+ax.set_xticks([0] + diameters + [6])
 #ax.set_yticks(yt)
-ax.set_xticklabels( ['0','1','','2','','3','','4','','5','6'], horizontalalignment='center')
+ax.set_xticklabels( ['0','1','2','3','4','5','6'], horizontalalignment='center')
 a=ax.axis([0,6,1,numpy.max(counts)*5])
 
 
@@ -72,5 +60,5 @@ a=ax.axis([0,6,1,numpy.max(counts)*5])
 
 
 
[…]

### 2013-06-11 split_plot.py

@@ -27,7 +27,7 @@ fig_size=(csize,csize*1.618)
 
 fig = pylab.figure()
 
-fig.subplots_adjust(top=0.95,bottom=0.1,left=0.10,right=0.93,hspace=0.35,wspace=0.2)
+fig.subplots_adjust(top=0.95,bottom=0.10,left=0.10,right=0.93,hspace=0.35,wspace=0.2)
 
 # this is plot 1 with Procut orange data
 
@@ -37,40 +37,49 @@ ax = fig.add_subplot(1,1,1)
 diameters = [ 1, 1.63, 2, 2.43, 3, 3.26, 4, 4.87 ]
 steps =[ '1', 'H', '2', 'J', '3', '2H', '4', '2J' ]
 counts = [ 3, 184, 409, 2445, 1060, 582, 93, 11 ] 
+types = [ '$S^{3}/Q$',  '$L(4,1)$' ,'$L(3,1)$' ,'$RP^{3}$' ,'$S^{3}$' ]
+markers = ['rs','ws','bx','ks','gs']
 
-diameters1= [ 1, 2, 3, 4, ]
-counts1= [ 3, 409, 1060, 93 ] 
+countd = {}
+countd[ '$S^{3}/Q$'] = [ 0, 1, 0, 0, 0, 0, 0, 0 ]
+countd[ '$L(4,1)$' ] = [ 0, 0, 0, 1, 0, 0, 0, 0 ]
+countd[ '$L(3,1)$' ] = [ 0, 0, 1, 1, 0, 0, 0, 0 ]
+countd[ '$RP^{3}$' ] = [ 0,10, 7, 5, 0, 0, 0, 0 ]
+countd[ '$S^{3}$'  ]  = [ 3, 173, 401, 2438, 1060, 582, 93, 11 ]
 
-diameters2 = [ 1.63, 2.43, 3.26, 4.87 ]
-steps2 =[ 'H', 'J', '2H', '2J' ]
-counts2 = [ 184, 2445, 582, 11 ] 
+#$1$ &  &  &  &   &3    & 3    \\
+#$2$ & 1&  &1 &17 &574  & 593  \\
+#$3$ &  &1 &1 &5  &3498 & 3505 \\
+#$4$ &  &  &  &   &675  & 675   \\
+#$5$ &  &  &  &   &11   & 11   \\
+#Total&1 &1 &2 &22   &4761   &4787  \\
+
+
+
 
 #ax.plot(diameters,counts,'k--',alpha=0.5)
[…]

### 2013-06-11 paper/paper.tex
combined figures and tables, tiny bit of writing

@@
 	\centering
-		\includegraphics[width=0.23\linewidth]{figures/hops.pdf}
-    \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$.}
+		\includegraphics[width=0.28\linewidth]{figures/hops.pdf}
+    \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$}
     	\label{fig:hop}
@@
 	\centering
-		\includegraphics[width=0.30\linewidth]{figures/jump.pdf}
+		\includegraphics[width=0.36\linewidth]{figures/jump.pdf}
     \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
@@
 Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric.
-Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:jump_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively.
+Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:hop_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively.
 We let $d(v,w)$ denote the \textbf{distance between vertices $v$ and $w$} obtained by minimizing over all paths containing edges as well as hops and jumps.
@@
 A small Python program was created to examine each manifold in the census and compute its diameter.
-See Figure \ref{fig:global_statistics} for a plot of the number of manifolds by topological type and diameter.
+See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter.
 As claimed, every manifold with diameter greater than $J$ is a 3-sphere.
-The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:global_statistics_edges}.
-Note that there are $3$-manifolds homeomor
+The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:type_statistics_edges}.
+Note the existence of $3$-manifolds with edge-diameter greater than half the maximum allowed edge-diameter and which are {\em not} homeomorphic to the 3-sphere.
+In particular, there are positively curved combinatorial 3-manifolds homeomorphic to $\mathbb{RP}^3$, $L(3,1)$ and $L(4,1)$ which have edge-diameter three.
 
@@
     \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
-    \end{center}
-    \caption{Distribution of diameters.}
-    \label{fig:global_statistics}
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/split_statistics.png}
-    \end{center}
-    \caption{Distribution of diameters for each type.}
-    \label{fig:type_statistics}
-\end{figure}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/edge_diameters.png}
-    \end{center}
-    \caption{Distribution of diameters when only edges are considered, without hops or jumps.}
-    \label{fig:global_statistics_edges}
-\end{figure}
-
-\begin{figure}
-\centering
+    {
+	    \setlength\fboxsep{10pt}
+		\setlength\fboxrule{0pt}
+		\fbox{\includegraphics[width=0.75\linewidth]{figures/split_statistics.png}}
+    }
     \begin{tabular} {| r | r | r | r | r | r | r |}
@@
 	\end{tabular}
-\caption{The positively curved combinatorial 3-manifolds by diameter and topological type}
-\label{fig:manifold_counts}
+    \end{center}
+    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below)}
+    \label{fig:type_statistics}
 \end{figure}
@@
 \begin{figure}
-\centering
+    \begin{center}
+    {
+	    \setlength\fboxsep{10pt}
+		\setlength\fboxrule{0pt}
+		\fbox{\includegraphics[width=0.75\linewidth]{figures/edge_diameters.png}}
+    }
 	\begin{tabular} {| r | r | r | r | r | r | r |}	
 		\hline
-Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
+		Edge-Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
 		\hline
@@
 	\end{tabular}
-\caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
-\label{fig:manifold_counts_edges}
+    \end{center}
+    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below)}
+	\label{fig:type_statistics_edges}
 \end{figure}
-
-\section{Algorithm Used}
-
-TO DO: BRIEF DISCUSSION OF ALGORITHM USED
 

### 2013-06-12 split_edge.py
updated split figures

@@ -27,16 +27,15 @@ fig_size=(csize,csize*1.618)
 
 fig = pylab.figure()
 
-fig.subplots_adjust(top=0.95,bottom=0.10,left=0.10,right=0.93,hspace=0.35,wspace=0.2)
+fig.subplots_adjust(top=0.95,bottom=0.1,left=0.10,right=0.93,hspace=0.35,wspace=0.2)
 
 # this is plot 1 with Procut orange data
 
 
 ax = fig.add_subplot(1,1,1)
 
-diameters = [ 1, 1.63, 2, 2.43, 3, 3.26, 4, 4.87 ]
-steps =[ '1', 'H', '2', 'J', '3', '2H', '4', '2J   ' ]
-counts = [ 3, 184, 409, 2445, 1060, 582, 93, 11 ] 
+diameters = [ 1, 2, 3, 4, 5 ]
+steps =[ '1', '2', '3', '4', '5' ]
 types = [ '$S^{3}/Q$',  '$L(4,1)$' ,'$L(3,1)$' ,'$RP^{3}$' ,'$S^{3}$' ]
 markers = ['ko','wo','bx','bo','ro']
 msdef = 9
@@ -44,37 +43,31 @@ markersizes = [msdef,msdef,12,msdef,msdef]
 mewdef = 1
 markeredgewidths = [mewdef,mewdef,1.1,mewdef,mewdef]
 
+
 countd = {}
-countd[ '$S^{3}/Q$'] = [ 0, 1, 0, 0, 0, 0, 0, 0 ]
-countd[ '$L(4,1)$' ] = [ 0, 0, 0, 1, 0, 0, 0, 0 ]
-countd[ '$L(3,1)$' ] = [ 0, 0, 1, 1, 0, 0, 0, 0 ]
-countd[ '$RP^{3}$' ] = [ 0,10, 7, 5, 0, 0, 0, 0 ]
-countd[ '$S^{3}$'  ]  = [ 3, 173, 401, 2438, 1060, 582, 93, 11 ]
-
-#$1$ &  &  &  &   &3    & 3    \\
-#$2$ & 1&  &1 &17 &574  & 593  \\
-#$3$ &  &1 &1 &5  &3498 & 3505 \\
-#$4$ &  &  &  &   &675  & 675   \\
-#$5$ &  &  &  &   &11   & 11   \\
[…]

### 2013-06-13 paper/paper.tex
updated split figures
bit of editing on abstract. worked on captions and other stuff for figures

@@
 \begin{abstract}
-The generalized sphere theorem of Grove and Shiohama is a central result in the differential geometry of positively curved manifolds.
+The generalized sphere theorem of Grove and Shiohama is an important result in the differential geometry of positively curved manifolds.
 It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere.
 In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds.
-That is, those for which at most five tetrahedra are incident at each edge.
-In the standard piecewise-linear metric, this bound is equivalent to requiring an angle deficit along each edge.
+That is, those with at most five tetrahedra incident at each edge, or equivalently, those with an angle deficit at each edge in the standard piecewise-linear metric.
 Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem.
-Here we prove a corresponding {\em discrete} generalized sphere theorem using a complete census of positively curved 3-manifolds completed by Lutz and Sullivan.
+Here we prove the corresponding {\em discrete} generalized sphere theorem using a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan.
 \end{abstract}
@@
     {
-	    \setlength\fboxsep{10pt}
+	    \setlength\fboxsep{15pt}
 		\setlength\fboxrule{0pt}
-		\fbox{\includegraphics[width=0.75\linewidth]{figures/split_statistics.png}}
+		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_statistics.png}}
     }
+    	\renewcommand{\arraystretch}{1.2}
     \begin{tabular} {| r | r | r | r | r | r | r |}
@@
     \end{center}
-    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below)}
+    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below).
+Note that all manifolds with diameter greater than half the maximum are 3-spheres.
+Results were computed using a 3-manifold census created by Lutz and Sullivan \cite{LutzSul_unpub}.}
     \label{fig:type_statistics}
@@
     {
-	    \setlength\fboxsep{10pt}
+	    \setlength\fboxsep{15pt}
 		\setlength\fboxrule{0pt}
-		\fbox{\includegraphics[width=0.75\linewidth]{figures/edge_diameters.png}}
+		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_statistics.png}}
     }
+    \renewcommand{\arraystretch}{1.2}
 	\begin{tabular} {| r | r | r | r | r | r | r |}	
@@
     \end{center}
-    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below)}
+    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below).
+Note that when diameters are measured in edges, there are manifolds with diameter greater than half the maximum which are {\em not} 3-spheres.
+Results were computed using a 3-manifold census created by Lutz and Sullivan \cite{LutzSul_unpub}.}
 	\label{fig:type_statistics_edges}

### 2013-06-13 paper/paper.tex
bit of editing on abstract. worked on captions and other stuff for figures

@@
 \noindent Interestingly, while the statement of Theorem \ref{thm:discrete_BM} uses edge-diameter, its proof relies on expanding the set of paths under consideration to include those which contain not only edges, but also other types of PL-paths between vertices.
+In this work, this enlarged set of paths will be crucial to obtaining a discrete result fully analogous to the classical generalized sphere theorem.
 
-\subsection{Hops and Jumps}
+\section{Hops and Jumps}
 The first new type of path is called a {\em hop}.
@@
 
-A small Python program was created to examine each manifold in the census and compute its diameter.
+A small Python program (available on request) was created to examine each manifold in the census and compute its diameter.
 See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter.

### 2013-06-13 paper/paper.tex

@@
 
-A small Python program (available on request) was created to examine each manifold in the census and compute its diameter.
+A small Python program (available on request) was created to examine each manifold in the census and compute its diameter and edge-diameter.
 See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter.
@@
 Note the existence of $3$-manifolds with edge-diameter greater than half the maximum allowed edge-diameter and which are {\em not} homeomorphic to the 3-sphere.
-In particular, there are positively curved combinatorial 3-manifolds homeomorphic to $\mathbb{RP}^3$, $L(3,1)$ and $L(4,1)$ which have edge-diameter three.
+However, we still have the following:
 
@@
 \section{Discussion}
-Why does adding additional paths above and beyond edge-paths restore the classical structure to our proof?
+In the classical version of the generalized sphere theorem, the diameter bound is sharp and exactly half the maximum diameter allowed by the classical Bonnet-Myers theorem.
+In the discrete setting discussed in this paper, we have seen that this relationship between diameter bounds cannot be maintained if distances are measured by edge-paths.
+However, if we expand the set of paths to include hops and jumps then we can prove a discrete generalized sphere theorem using a sharp diameter bound exactly half that of our discrete Bonnet-Myers theorem.
 

### 2013-06-14 paper/paper.tex
more writing, fixed some small problems with the graphs

--- 2013-06-13 paper/paper.tex
@@
 Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem.
-Here we prove the corresponding {\em discrete} generalized sphere theorem using a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan.
+Here, we use a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan to prove a corresponding {\em discrete} generalized sphere theorem.
 \end{abstract}
@@
 Similarly, we let $diam(M)$ denote the \textbf{diameter of $M$} defined in terms of the distance function $d$.
-Let $\mathcal{P}$ denote the set of all the paths containing edges, hops or jumps.
+Finally, let $\mathcal{P}$ denote the set of all the paths containing edges, hops or jumps.
 
@@
 If we naively imitate the classical results in the discrete setting, from Theorem \ref{thm:discrete_BM} we would want to assume the bound $diam_1(M)>\frac{5}{2}$.
-Unfortunately, there are positively curved combinatorial 3-manifolds with edge-diameter three that are not homeomorphic to the 3-sphere.
-In particular, as we will see, $\mathbb{RP}^3$ has positively curved triangulations with edge-diameter three.
-However, if we instead use the finer-grained measure of diameter, $diam(M)$, the problem disappears and the discrete results mirror the classical ones exactly.
+Right away we see a discrepancy between the discrete and classical results since this bound cannot possibly be sharp.
+Worse, the smallest weakening that could be sharp is $diam_1(M) \geq 3$ and this bound does not work.
+As we will see, there are positively curved combinatorial 3-manifolds $M$ with $diam_1(M)=3$ but which are {\em not} homeomorphic to the 3-sphere.
+However, if we instead use the finer-grained measure of diameter, $diam(M)$, the problem disappears and the discrete results mirror the classical smooth ones exactly.
 That is, we can prove a discrete version of the generalized sphere theorem with a diameter bound exactly half the maximum allowed by the corresponding discrete Bonnet-Myers theorem.
@@
 
-\begin{thm} Any connected positively curved combinatorial 3-manifold $M$ with $diam(M)>J$ is homeomorphic to the 3-sphere.
-This bound is sharp, and equal to half the maximum diameter which occurs for any such $M$.
+\begin{thm} Let $M$ be a connected, positively curved, combinatorial 3-manifold.
+If $M$ has $diam(M)>J$ then it is homeomorphic to a 3-sphere.
+Moreover, this bound is sharp and equal to half the maximum diameter which occurs for any such $M$.
 \label{thm:discrete_GS}
@@
 
-\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types, see \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}.
+\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types.
+See \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}.
 This census contains 4787 manifolds, with the following topologies represented: the 3-sphere $S^3$, the real projective space $\mathbb{RP}^3$, the lens spaces $L(3,1)$ and $L(4,1)$, and the cube space $S^3/Q$.
@@
 Note the existence of $3$-manifolds with edge-diameter greater than half the maximum allowed edge-diameter and which are {\em not} homeomorphic to the 3-sphere.
-In particular, there are positively curved combinatorial 3-manifolds homeomorphic to $\mathbb{RP}^3$, $L(3,1)$ and $L(4,1)$ which have edge-diameter three.
+
+If one absolutely insists on results stated in terms of edge-diameter then the census does give us the following result:
+
+\begin{thm} If $M$ is a connected, positively curved, combinatorial 3-manifold satisfying the bound $diam_1(M) > 3$ then $M$ is homeomorphic to the 3-sphere. Moreover, this bound is sharp.
+\end{thm}
 
@@
 		\setlength\fboxrule{0pt}
-		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_statistics.png}}
+		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_edge_diameters.png}}
     }
@@
 \section{Discussion}
-Why does adding additional paths above and beyond edge-paths restore the classical structure to our proof?
-
 We should mention another notable advantage of expanding the space of paths to include hops and edges.

### 2013-06-14 split_edge.py

@@ -46,8 +46,8 @@ markeredgewidths = [mewdef,mewdef,1.1,mewdef,mewdef]
 
 countd = {}
 countd[ '$S^{3}/Q$'] = [ 0, 1, 0, 0, 0]
-countd[ '$L(4,1)$' ] = [ 0, 0, 0, 1, 0]
-countd[ '$L(3,1)$' ] = [ 0, 0, 1, 1, 0]
+countd[ '$L(4,1)$' ] = [ 0, 0, 1, 0, 0]
+countd[ '$L(3,1)$' ] = [ 0, 1, 1, 0, 0]
 countd[ '$RP^{3}$' ] = [ 0,17, 5, 0, 0]
 countd[ '$S^{3}$'  ] = [ 3, 574, 3498, 675, 11 ]
 
@@ -60,7 +60,7 @@ for t,m,ms,mew in zip(types,markers,markersizes,markeredgewidths):
 
     ax.semilogy(diameters,countd[t],m,ms=ms,mew=mew,label=t)
 
-ax.set_xlabel('Diameter with only Edges')
+ax.set_xlabel('Edge-Diameter')
 #ax.xaxis.set_label_coords(0.5, -0.12)
 ax.set_ylabel('Count (log scale)')
 ax.set_xticks([0] + diameters + [6])

### 2013-06-14 paper/paper.tex
more writing

@@
 We should mention another notable advantage of expanding the space of paths to include hops and edges.
-In classical differential geometry, the starting point, the initial direction, and length completely determine a geodesic.
-There is no branching of geodesics.
-This property of fundamental importance to many proofs in Riemannian geometry.
-If we use only edge-paths, then in a positi`vely curved combinatorial 3-manifold it is possible to have two minimal paths with the same length and which begin by traversing the same edge, yet reach distinct destination vertices.
-That is, using only edge paths, minimal geodesics can ``branch''.
+In classical differential geometry, a geodesic segment is completely determined by its the starting point, initial direction, and length.
+There is no ``branching'' of geodesics on a Riemannian manifold.
+Indeed, this property is of fundamental importance to constructing many arguments in this area.
+However in the discrete setting, if we use only edge-paths then this valuable characteristic does not hold.
+It is possible to have two minimal edge-paths with the same length and which begin by traversing the same edge, yet reach distinct destination vertices.
 
-This important property is restored, at least for positively curved manifolds, if we include hops and jumps. In \cite{Trout10} the following is shown.
+To see this, consider a combinatorial 3-manifold $M$ containing a minimal jump from vertex $v$ to vertex $w$ passing through edges $e_1$ and $e_2$ (as seen in Figure \ref{fig:jump}.) Using the structure of a jump, we will construct two edge-paths starting at $v$.
+We choose the first edge from $v$ to some vertex in $e_1$ and then we can continue the path to either of the vertices in $e_2$.
+The resulting paths, $Q_1$ and $Q_2$, must be minimal among edge-paths.
+If they were not, then their endpoints would be connected by an edge and a two-edge path from $v$ to $w$ would exist, contradicting the minimality of the jump.
+Thus $Q_1$ and $Q_2$ provide examples of minimal edge-paths which ``branch''.
+
+The non-branching property of minimal paths is restored, at least for positively curved manifolds, if we include hops and jumps. See \cite{Trout10}. In particular we have:
 
@@
 
+\noindent It is not surprising that the
+
 \bibliographystyle{plain}

### 2013-06-14 paper/paper.tex
more writing. first draft done

@@
 As we will see, there are positively curved combinatorial 3-manifolds $M$ with $diam_1(M)=3$ but which are {\em not} homeomorphic to the 3-sphere.
-However, if we instead use the finer-grained measure of diameter, $diam(M)$, the problem disappears and the discrete results mirror the classical smooth ones exactly.
+However, if we instead use the finer-grained measure of diameter, $diam(M)$, the problem disappears and the discrete results mirror the classical ones exactly.
 That is, we can prove a discrete version of the generalized sphere theorem with a diameter bound exactly half the maximum allowed by the corresponding discrete Bonnet-Myers theorem.
@@
 Thus $Q_1$ and $Q_2$ provide examples of minimal edge-paths which ``branch''.
-
-The non-branching property of minimal paths is restored, at least for positively curved manifolds, if we include hops and jumps. See \cite{Trout10}. In particular we have:
+The non-branching property of minimal paths is restored, at least for positively curved manifolds, if we include hops and jumps.
+In particular we have:
 
@@
 
-\noindent It is not surprising that the
+\noindent See \cite{Trout10} for details.
+
+It is not entirely surprising that discrete differential geometry on a PL-manifold $M$ comes to resemble classical differential geometry more and more as the set of admissible PL-paths on $M$ is expanded.
+Forcing ourselves to use only edges amounts to coarse-graining the underlying PL-geometry, and expanding the set of paths clearly reduces this coarse-graining.
+We could even include {\em all} the PL-paths in $M$ and still have a ``discrete'' geometry in some sense, because our PL-metric is determined solely by the structure of $M$ as a simplicial complex.
+However, since PL-paths admit infinitesimal deformations we would lose much of the ``discrete'' flavor to the theory.
+The successful use of hops and jumps in this paper and in \cite{Trout10} indicate that one can improve the behavior of discrete geometries by judicious inclusion of PL-paths in addition to edges.
+Investigating the possibilities for such inclusions and their benefits represents an exciting topic of further research.
 

### 2013-06-18 paper/paper.tex
added code to generate table. edited paper.

@@
 \usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
+\usepackage{deluxetable}             
 \usepackage{graphicx}              % to include figures
@@
 Differential geometry is of central importance not only to geometers but also topologists, physicists and, increasingly, many of those interested in applied topics such as finite element analysis and computer graphics.
-One new offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogues to the classical theorems and concepts from differential geometry.
+One significant offshoot in this area is {\em discrete differential geometry} (DDG) which seeks discrete analogues to the classical theorems and concepts from differential geometry.
 Since many computational treatments of differential geometry involve discretizing shapes, this subject has particular relevance for those with applied interests, see \cite{grinspun2006discrete} for examples.
@@
 A particularly important goal in classical differential geometry is to elucidate the relationship between the curvature of a Riemannian (or semi-Riemannian) space and its topology. The classical
-results in this area are numerous, beautiful, and have inspired an
+results in this area are numerous, elegant, and have inspired an
 enormous amount of subsequent research. In this paper, we will present a discrete analogue of the ``generalized'' sphere theorem of Grove and Shiohama \cite{groveshiohama}.
@@
 
-\noindent The proof of this theorem relies on the remarkable work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types.
+\noindent The proof of this theorem relies on the work of Lutz, Sullivan and Sulanke in \cite{Lutz07, LutzSul_unpub, sulanke2009isomorphism} who created a complete census of positively curved combinatorial 3-manifolds along with a list of their topological types.
 See \cite{Lutz_online_manifolds} and \cite{Lutz_online_topological_types}.
@@
 
-A small Python program (available on request) was created to examine each manifold in the census and compute its diameter and edge-diameter.
+Python programs (available on request) were written to examine each manifold in the census and compute its diameter and edge-diameter.
 See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter.
@@
 
-If one absolutely insists on results stated in terms of edge-diameter then the census does give us the following result:
+When only edge-paths are considered, as expected we obtain the following result from the census:
 
@@
 Investigating the possibilities for such inclusions and their benefits represents an exciting topic of further research.
-
 \bibliographystyle{plain}
 \bibliography{manifolds}
+
+\input{bigtable.tex}
+
 \end{document}

### 2013-06-18 maketable.py

@@ -1,0 +1,43 @@
+import Table
+
+fin = open('output.txt','r')
+fout = open('./paper/bigtable.tex','w')
+
+all = fin.readlines()
+fin.close()
+for i in range(len(all)):
+    all[i] = all[i].replace(' ','$',1).replace(' ','$ $',1).replace('$',' $',1).replace('\n','$\n')
+    all[i] = all[i].replace('E','')
+    all[i] = all[i].replace('1H','H')
+    all[i] = all[i].replace('1J','J')
+    print all[i].split()
+    print '\n'
+
+max = len(all)
+N = 5
+extra = 1
+
+chunks = [all[x:x+max/(N)+extra] for x in xrange(0,len(all),max/(N)+extra)]
+chunks[-1] = chunks[-1] + ['']*(len(chunks[-2])-len(chunks[-1]))
+for c in chunks:
+    print len(c)
+
+print len(chunks)
+
+
+t = Table.Table(N, justs='l'*N, caption='Awesome results', label="tab:full")
+t.add_header_row(['HEAD']*N)
+
+t.add_data(chunks)
+t.print_table(fout)
+fout.close()
+gin = open('./paper/bigtable.tex','r')
+raw = gin.readlines()
+gin.close()
+out = ''
+fout = open('./paper/bigtable.tex','w')
+
+head = '$N$ & $t$ & $d$ &' * N
+for i in raw:
+    fout.write(i.replace(r'\begin{deluxetable}{'+'l'*N+'}',r'\begin{deluxetable}{'+'|lll|'*N+'}').replace('$ $','$ & $').replace(r'\tablecolumns{'+str(N)+'}',r'\tablecolumns{'+str(N*3)+'}').replace(' $',' & $',1).replace(r'\colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD}',head[:-1] ))
+fout.close()

### 2013-06-18 maketable.py
changed table alignment

@@ -39,5 +39,5 @@ fout = open('./paper/bigtable.tex','w')
 
 head = '$N$ & $t$ & $d$ &' * N
 for i in raw:
-    fout.write(i.replace(r'\begin{deluxetable}{'+'l'*N+'}',r'\begin{deluxetable}{'+'|lll|'*N+'}').replace('$ $','$ & $').replace(r'\tablecolumns{'+str(N)+'}',r'\tablecolumns{'+str(N*3)+'}').replace(' $',' & $',1).replace(r'\colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD}',head[:-1] ))
+    fout.write(i.replace(r'\begin{deluxetable}{'+'l'*N+'}',r'\begin{deluxetable}{'+'|llr|'*N+'}').replace('$ $','$ & $').replace(r'\tablecolumns{'+str(N)+'}',r'\tablecolumns{'+str(N*3)+'}').replace(' $',' & $',1).replace(r'\colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD} & \colhead{HEAD}',head[:-1] ))
 fout.close()

### 2013-06-25 split_edge.py
messed with tables

@@ -5,9 +5,9 @@ from matplotlib.font_manager import fontManager, FontProperties
 #set fonts here
 font= FontProperties(size='small')
 params = {'backend': 'ps',
-          'axes.labelsize': 18,
-          'xtick.labelsize': 16,
-          'ytick.labelsize': 16,
+          'axes.labelsize': 16,
+          'xtick.labelsize': 15,
+          'ytick.labelsize': 15,
           'legend.fontsize': 16,
           #'font.family': 'sans',
           'font.serif': ['Times','Palatino','serif'],
@@ -37,7 +37,7 @@ ax = fig.add_subplot(1,1,1)
 diameters = [ 1, 2, 3, 4, 5 ]
 steps =[ '1', '2', '3', '4', '5' ]
 types = [ '$S^{3}/Q$',  '$L(4,1)$' ,'$L(3,1)$' ,'$RP^{3}$' ,'$S^{3}$' ]
-markers = ['ko','wo','bx','bo','ro']
+markers = ['ws','wo','bx','bs','ro']
 msdef = 9
 markersizes = [msdef,msdef,12,msdef,msdef]
 mewdef = 1

### 2013-06-25 paper/paper.tex
messed with tables
journal specific stuff added

@@
+% EJC papers *must* begin with the following two lines. 
 \documentclass[12pt]{article}
+\usepackage{e-jc}
 
-\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
-\usepackage{deluxetable}             
-\usepackage{graphicx}              % to include figures
-\usepackage{amsmath}               % great math stuff
-\usepackage{amsfonts}              % for blackboard bold, etc
-\usepackage{amsthm}                % better theorem environments
-\usepackage{url}
+% Please remove all other commands that change parameters such as
+% margins or pagesizes.
 
-% various theorems, numbered by section
+% only use standard LaTeX packages
+% only include packages that you actually need
 
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{conj}[thm]{Conjecture}
-\newtheorem{dfn}[thm]{Definition}
+% we recommend these ams packages
+\usepackage{amsthm,amsmath,amssymb}
 
-\DeclareMathOperator{\id}{id}
+% we recommend the graphicx package for importing figures
+\usepackage{graphicx}
 
-\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
-\newcommand{\RR}{\mathbb{R}}      % for Real numbers
-\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
-\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
-\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
+% use this command to create hyperlinks (optional and recommended)
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
+
+% use these commands for typesetting doi and arXiv references in the bibliography
+\newcommand{\doi}[1]{\href{http://dx.doi.org/#1}{\texttt{doi:#1}}}
+\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}
+
+% all overfull boxes must be fixed; 
+% i.e. there must be no text protruding into the margins
+
+% declare theorem-like environments
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{fact}[theorem]{Fact}
+\newtheorem{observation}[theorem]{Observation}
+\newtheorem{claim}[theorem]{Claim}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\newtheorem{open}[theorem]{Open Problem}
+\newtheorem{problem}[theorem]{Problem}
+\newtheorem{question}[theorem]{Question}
+
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
+\newtheorem{note}[theorem]{Note}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and diameter greater than $\frac{\pi}{2\sqrt{\delta}}$.
+\begin{theorem}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and diameter greater than $\frac{\pi}{2\sqrt{\delta}}$.
 Then, $M$ is homeomorphic to a sphere.
 \label{thm:grove_shiohama}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with sectional curvature $K \geq \delta > 0$.
+\begin{theorem}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with sectional curvature $K \geq \delta > 0$.
 Then the diameter of $M$ is at most $\frac{\pi}{\sqrt{\delta}}$.
 \label{thm:bonnet_myers}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{dfn}The \textbf{edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length (as a PL-path in the standard PL-metric) of an edge-path from $v$ to $w$.
+\begin{definition}The \textbf{edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length (as a PL-path in the standard PL-metric) of an edge-path from $v$ to $w$.
 We denote this quantity $d_1(v,w)$.
 The \textbf{edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$.
-\end{dfn}
+\end{definition}
 
@@
 
-\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
+\begin{theorem}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
 \label{thm:discrete_BM}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ are vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
+\begin{definition}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ are vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$.
 The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a \textbf{hop} from $v_1$ to $v_2$.
 See Figure \ref{fig:hop} for an illustration.
-\end{dfn}
+\end{definition}
 
@@
 
-\begin{dfn}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$.
+\begin{definition}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$.
 We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a \textbf{jump} from $v_1$ to $v_2$.
 For an illustration, see Figure \ref{fig:jump}.
-\end{dfn}
+\end{definition}
 
@@
 
-\begin{lem}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
-\end{lem}
+\begin{lemma}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
+\end{lemma}
 
@@
 
-\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
+\begin{theorem}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
 \label{thm:discrete_BM_expanded_paths}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{thm} Let $M$ be a connected, positively curved, combinatorial 3-manifold.
+\begin{theorem} Let $M$ be a connected, positively curved, combinatorial 3-manifold.
 If $M$ has $diam(M)>J$ then it is homeomorphic to a 3-sphere.
@@
 \label{thm:discrete_GS}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{thm} If $M$ is a connected, positively curved, combinatorial 3-manifold satisfying the bound $diam_1(M) > 3$ then $M$ is homeomorphic to the 3-sphere. Moreover, this bound is sharp.
-\end{thm}
+\begin{theorem} If $M$ is a connected, positively curved, combinatorial 3-manifold satisfying the bound $diam_1(M) > 3$ then $M$ is homeomorphic to the 3-sphere. Moreover, this bound is sharp.
+\end{theorem}
 
@@
 
-\begin{thm}(Unique Extension) If two minimal paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of a positively curved 3-manifold $M$ and $P_1$ and $P_2$ have the same length, then $P_1 = P_2$.
+\begin{theorem}(Unique Extension) If two minimal paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of a positively curved 3-manifold $M$ and $P_1$ and $P_2$ have the same length, then $P_1 = P_2$.
 This result is false if $\mathcal{P}$ is replaced by $\mathcal{P}_1$.
-\end{thm}
+\end{theorem}
 
@@
 
-\input{bigtable.tex}
+% \input{bigtable.tex}
 

### 2013-06-25 paper/paper.tex
more journal specific stuff added

@@
 
+\title{A Generalized Sphere Theorem for\\Positively Curved Combinatorial 3-Manifolds}
+
+\author{Aaron D. Trout\\
+\small Department of Mathematics\\[-0.8ex]
+\small Chatham University\\[-0.8ex] 
+\small Pittsburgh PA 15232,  U.S.A.\\
+\small\tt atrout@chatham.edu\\
+\and
+Vadas Gintautas\\
+\small Department of Physics\\[-0.8ex]
+\small Chatham University\\[-0.8ex] 
+\small Pittsburgh PA 15232,  U.S.A.\\
+\small\tt vgintautas@chatham.edu}
+
+\date{\dateline{June 25, 2013}{XX}\\
+\small Mathematics Subject Classifications: 52C99, 53A99, 57M99}
+
 \begin{document}
-
-\nocite{*}
-
-\title{A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds}
-
-\author{Aaron Trout \\ Department of Mathematics \\
-Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA \and
-Vadas Gintautas\\ Department of Physics \\
-Chatham University \\ Woodland Rd, Pittsburgh PA 15232 USA}
 
@@
 Here, we use a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan to prove a corresponding {\em discrete} generalized sphere theorem.
+
+\bigskip\noindent \textbf{Keywords:} discrete differential geometry; generalized sphere theorem; piecewise-linear geometry; positively curved combinatorial manifold
 \end{abstract}

### 2013-06-26 paper/supplemental.tex
moved big table to separate file

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{deluxetable}             
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{url}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{dfn}[thm]{Definition}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\begin{document}

\title{A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds: Supplemental Information}

\maketitle

\section{Table of Diameters}

The following table lists the diameter of each of the positively curved 3-manifolds.  Here $N$ refers to the manifold number as tabulated by Lutz, Sullivan and Sulankein (see citations in paper), $t$ refers to the topological type, and $d$ refers to diameter, when hops ($H = \frac{2}{3}\sqrt{2}$) and jumps ($J = \frac{1}{2}\sqrt{2} + \sqrt{3}$) are considered.
\input{bigtable.tex}

\end{document}

### 2013-06-26 maketable.py

@@ -25,7 +25,7 @@ for c in chunks:
 print len(chunks)
 
 
-t = Table.Table(N, justs='l'*N, caption='Awesome results', label="tab:full")
+t = Table.Table(N, justs='l'*N, caption='Diameters', label="tab:full")
 t.add_header_row(['HEAD']*N)
 
 t.add_data(chunks)

### 2013-06-27 paper/paper.tex
added link to github material in bibliography and referenced it in paper

@@
 
-Python programs (available on request) were written to examine each manifold in the census and compute its diameter and edge-diameter.
+Python programs were written to examine each manifold in the census and compute its diameter and edge-diameter.
+Code is available online \cite{Gintautas_Trout_github_manifold} along with a complete table of results for each manifold.
 See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter.
@@
 
-% \input{bigtable.tex}
-
 \end{document}
